---
title: "Consolidate Western Provisional Seed Zones"
author: "steppe"
date: "2023-09-14"
output: html_document
---
We will also consolidate the provisional seed zones. 
The provisional seed zones are used more commonly than the empirical, because most species
have not yet had their empirical zones produced. 

```{r}
library(tidyverse)
library(sf)
library(terra)
```

Currently, these data exist as a mix of vector and raster data. 
We will consolidate them into a raster stack, which can be readily queried, and requires
less storage and active memory. 

```{r load data}
p <- '../data/geodata/raw_provisional'
f <- file.path(p, list.files(p, recursive = T, pattern = 'tif$|shp$'))

Bower2013 <- st_read(f[1], quiet = TRUE) %>% 
  select(seed_zone) 

GreatBasin_Bower <- st_read(f[2], quiet = TRUE) %>% 
  select(seed_zone)

DesertSW <- st_read(f[3], quiet = TRUE) %>% 
  select(seed_zone = Zone) 

Mojave <- st_read(f[5], quiet = TRUE) %>% 
  rename(seed_zone = Zone) 
```

Now convert these data into raster format.  

```{r, eval = F}

domain <- st_bbox(c(xmin = -125.1, xmax = -100, ymax = 52.1, ymin = 32.0), crs = st_crs(4326))
template <- terra::rast(x = ext(domain), crs = "epsg:4326", res = 1/360) # 10 arc second

plot(ext(rs))

#' combine the provisional seed zones as rasters
#' @param x sf/dataframe/tibble to transform to a raster
#' @param bbox the domain to crop data to, optional
#' @param template raster template with the appropriate domain, grain, and crs to convert x to. 
#' @param outpath an output directory to save the raster
prov_writer <- function(x, bbox, template, outpath){

  ob_name <- deparse(substitute(x))
  if(any(sf::st_is_valid(x) == FALSE)) {x <- sf::st_make_valid(x)}
  if(terra::crs(x) != terra::crs(template)){x <- sf::st_transform(x, terra::crs(template))}
  if(!missing(bbox)){x <- sf::st_crop(x, sf::st_bbox(template))}
  
  x_r <- terra::rasterize(x, template, field = 'seed_zone', 
                          filename = file.path(outpath, paste0(ob_name, '.tif')))
}

prov_writer(Bower2013, bbox = domain, template = template, outpath = '../data/geodata/provisional_products')
prov_writer(DesertSW, bbox = domain, template = template, outpath = '../data/geodata/provisional_products')
prov_writer(GreatBasin_Bower, bbox = domain, template = template, outpath = '../data/geodata/provisional_products')
prov_writer(Mojave, bbox = domain, template = template, outpath = '../data/geodata/provisional_products')

rm(prov_writer, template, domain, f, p, Bower2013, DesertSW, GreatBasin_Bower, Mojave)
```

now combine these raster layers into a single stack

```{r}
p <- '../data/geodata/provisional_products'
f <- file.path(p, list.files(p, pattern = 'tif$'))
r <- rast(f)
names(r) <- gsub('\\.tif', '', basename(sources(r)))

writeRaster(r, filename = '../data/geodata/final_provisional/provisional_stack.tif')
rm(p, f, r)
getwd()
```

