---
title: "Consolidate Western Empirical Seed Zones"
author: "steppe"
date: "2023-09-14"
output: html_document
---

Roughly 2 dozen western species have empirical seed transfer zones and climate matched seed zones. 
However, nearly all of the spatial vector products derived from these experiments are in variable formats.
Here we attempt to make the naming and contents of this projects unified. 
Further in the end we produce a raster stack of all products to enable the rapid retrieval of empirical seed zone information. 

We first attempted to use wget to download the links, however the webpage is difficult... 
```{sh List all available Seed Zones, eval = F}
# sudo apt install lynx -y
lynx -listonly -dump https://www.fs.usda.gov/wwetac/threat-map/TRMSeedZoneData.php | awk '/http/{print $2}' | grep -E '\.zip' > ../data/raw/links.txt
# manually removed the three records for provisional zones
cd ../data/raw
ls
#wget -b links.txt -P ../geodata/raw/
```


```{r Create all tables, eval = F}

data.frame(
  Species = c(
    'Astragalus lonchocarpus', 'Bouteloua gracilis', 'Heliomeris multiflora',
    'Koeleria macrantha', 'Elymus elymoides', 
    
    'Elymus glaucus', 'Elymus elymoides', 'Leymus cinereus',
    'Bromus carinatus',  'Koeleria macrantha', 'Pseudoroegneria spicata', 'Poa secunda', 
    'Allium acuminatum', 'Achnatherum hymenoides', 'Holodiscus discolor', 
    'Achnatherum thurberianum', 'Artemisia tridentata spp. vaseyana', 'Artemisia tridentata spp. wyomingensis', 
    'Artemisia tridentata ssp. tridentata',
    
    'Sphaeralcea ambigua', 'Ephedra nevadensis', 
    'Macheranthera canescens', 'Cleome lutea', 'Sphaeralcea parvifolia', 
    'Sporobolus cryptandrus', 'Pleuraphis jamesii'
    ), 
  PLANTS_code = c(
    'ASLO3', 'BOGR2', 'HEMU3',
    'KOMA', 'ELEL5',
    
    'ELGL', 'ELEL5', 'LECI4',
    'BRCA5', 'KOMA', 'PSSP6',
    'POSE', 'ALAC4', 'ACHY',
    'HODI', 'ACHT7', 'ARTRV',
    'ARTRW8', 'ARTRT', 
    
    'SPAM2', 'EPNE',
    'MACA3', 'CLLU2', 'SPPA2',
    'SPCR', 'PLJA'
    ), 
  Method_broad = c(
    rep('Climate-Matched', times = 5),
    rep('Empirical', times = 21)
  ),
  Method_fine = c(
    rep('Climate-Matched', times = 5),
    rep('Common Garden', times = 14),
    rep('Landscape Genetic', times = 7)
    )
) |> 
  write.csv('../data/raw/STZ_Methods.csv', row.names = F)

data.frame(
  PLANTS_code = c(
    'ASLO3', 'BOGR2', 'HEMU3',
    'KOMA', 'ELEL5',
    
    'ELGL', 'ELEL5', 'LECI4',
    'BRCA5', 'KOMA', 'PSSP6',
    'POSE', 'ALAC4', 'ACHY',
    'HODI', 'ACHT7', 'ARTRV',
    'ARTRW8', 'ARTRT', 
    
    'SPAM2', 'EPNE',
    'MACA2', 'CLLU2', 'SPPA2',
    'SPCR', 'PLJA'
    ),
  Range = c(
    rep('Colorado Plateau', times = 5),
    'Blue Mountains', 'Western US', 'Columbia Basin - Great Basin',
    'Blue Mountains', 'Columbia Basin - Great Basin',
    rep('Western US', times = 9), rep('Mojave Desert', times = 2),
    rep('Colorado Plateau', times = 2), 'Colorado Plateau - Great Basin',
    'Colorado Plateau', 'Colorado Plateau - Great Basin - Mojave Desert - Western Plains'
  ),
  Range_Code = c(
    rep('CP', times = 5),
    'BM', 'WUS', 'CB - GB', 
    'BM',  'CB - GB', rep('WUS', times = 9), rep('MD', times = 2),
    rep('CP', times = 2), 'CP - GB',
    'CP', 'CP - GB - MD - WP'
    )
) |> 
  write.csv('../data/raw/Spatial_Domains_STZ.csv', row.names = F)

data.frame(
  PLANTS_code = c(
    'ASLO3', 'BOGR2', 'HEMU3',
    'KOMA', 'ELEL5',
    
    'ELGL', 'ELEL5', 'LECI4',
    'BRCA5', 'KOMA', 'PSSP6',
    'POSE', 'ALAC4', 'ACHY',
    'HODI', 'ACHT7', 'ARTRV',
    'ARTRW8', 'ARTRT', 
    
    'SPAM2', 'EPNE',
    'MACA2', 'CLLU2', 'SPPA2',
    'SPCR', 'PLJA'
    ),
  Range_Code = c(
    rep('CP', times = 5),
    'BM', 'WUS', 'CB - GB', 
    'BM',  'CB - GB', rep('WUS', times = 9), rep('MD', times = 2),
    rep('CP', times = 2), 'CP - GB',
    'CP', 'CP - GB - MD - WP'
    ),
  Bibtex_key = c(
    rep('doherty2017matching', times = 5),
    'erickson2004landscape', 'UNKNOWN', 'johnson2016linking',
    'johnson2010mapping', 'UNKNOWN', 'bradley2013genetic',
    'johnson2015relating', 'johnson2013genecology', 'johnson2012genecology', 
    'horning2010genecology', 'johnson2017genecology', rep('UNKNOWN', times = 3),
    rep('shryock2017landscape', times = 2), 
    rep('massatti2020genetically', times = 2), 
    rep('massatti2019genetically', times = 3)
  )
) |> 
  write.csv('../data/raw/Citations.csv', row.names = F)

data.frame(
    USDA_Species = c(
    'Astragalus lonchocarpus', 'Bouteloua gracilis', 'Heliomeris multiflora',
    'Koeleria macrantha', 'Elymus elymoides', 
    
    'Elymus glaucus', 'Elymus elymoides', 'Leymus cinereus',
    'Bromus carinatus',  'Koeleria macrantha', 'Pseudoroegneria spicata', 'Poa secunda', 
    'Allium acuminatum', 'Achnatherum hymenoides', 'Holodiscus discolor', 
    'Achnatherum thurberianum', 'Artemisia tridentata spp. vaseyana', 'Artemisia tridentata spp. wyomingensis', 
    'Artemisia tridentata ssp. tridentata',
    
    'Sphaeralcea ambigua', 'Ephedra nevadensis', 
    'Macheranthera canescens', 'Cleome lutea', 'Sphaeralcea parvifolia', 
    'Sporobolus cryptandrus', 'Pleuraphis jamesii'
    ), 
    Species = c(
    'Astragalus lonchocarpus', 'Bouteloua gracilis', 'Heliomeris multiflora',
    'Koeleria macrantha', 'Elymus elymoides', 
    
    'Elymus glaucus', 'Elymus elymoides', 'Leymus cinereus',
    'Bromus sitchensis ssp. carinatus',  'Koeleria macrantha', 'Pseudoroegneria spicata', 'Poa secunda', 
    'Allium acuminatum', 'Eriocoma hymenoides', 'Holodiscus discolor', 
    'Eriocoma thurberiana', 'Artemisia tridentata spp. vaseyana', 'Artemisia tridentata spp. wyomingensis', 
    'Artemisia tridentata ssp. tridentata',
    
    'Sphaeralcea ambigua', 'Ephedra nevadensis', 
    'Macheranthera canescens', 'Cleomella lutea', 'Sphaeralcea parvifolia', 
    'Sporobolus cryptandrus', 'Pleuraphis jamesii'
    ), 
  PLANTS_code = c(
    'ASLO3', 'BOGR2', 'HEMU3',
    'KOMA', 'ELEL5',
    
    'ELGL', 'ELEL5', 'LECI4',
    'BRCA5', 'KOMA', 'PSSP6',
    'POSE', 'ALAC4', 'ACHY',
    'HODI', 'ACHT7', 'ARTRV',
    'ARTRW8', 'ARTRT', 
    
    'SPAM2', 'EPNE',
    'MACA2', 'CLLU2', 'SPPA2',
    'SPCR', 'PLJA'
    ) 
)  |> 
  write.csv('../data/raw/Taxonomy.csv', row.names = F)


```

```{r}
library(tidyverse)
library(sf)
```

The files have no consistency in naming. We will provide a similar format to name all files. 

```{r Rename all geodata, eval = F}

p2raw <- '../data/geodata/raw/'

raw_names <- data.frame(
  Path_Name = paste0(p2raw, list.files(p2raw)),
  Raw_Name = list.files(p2raw),
  New_Name = c(
    'Eriocoma_hymenoides-WUS', 'Allium_acuminatum-WUS', 'Artemisia_tridentata-WUS', 
    'Artemisia_tridentata_spp._vaseyana-WUS', 'Astragalus_lonchocarpus-CP', 'Astragalus_lonchocarpus-GSTZ', 
    'Pseudoroegneria_spicata-WUS', 'Bouteloua_gracilis-CP', 'Cleome_lutea-CP', 'Cleome_serrulata-', 
    'Elymus_elymoides-CP', 'Elymus_elymoides-WUS', 'Elymus_glaucus-BM', 
    'Ephedra_nevadensis-MD', 'Heliomeris_multiflora-CP', 
    'Koeleria_macrantha-CB-GB', 'Leymus_cinereus-CB-GB', 'Machaeranthera_canescens-CP', 
    'Bromus_sitchensis_ssp._carinatus-BM', 'Pleuraphis_jamesii-CP-GB-MD-WP', 
    'Poa_secunda-WUS', 'Sporobolus_cryptandrus-CP', 'Sphaeralcea_ambigua-MD',
    'Sphaeralcea_parvifolia-CP-GB'
    )
)
  
raw_names2 <- paste0(p2raw, list.files(p2raw)) #
fnames <- lapply(raw_names2, unzip, list = T) # need individual file names in the compressed directory
names(fnames) <- list.files(p2raw) # we will add the zipped directory as an identified to the files
raw_names2 <- data.table::rbindlist(fnames, idcol = 'ZipName') |>
  dplyr::select(-Length, -Date) 
raw_names2 <- dplyr::left_join(raw_names2, raw_names, by = c('ZipName' = 'Raw_Name'))

lapply(unique(raw_names2$Path_Name), unzip, junkpaths = T, exdir = '../data/geodata/step1')

raw_names2 <- raw_names2 %>% 
  mutate(
    New_Name = paste0(New_Name,  stringr::str_extract(Name, '[.].*')),
    Name = sub('.*\\/', "", Name)
  ) %>% 
  filter(str_detect(Name, '[.]'))

file.rename(from = paste0('../data/geodata/step1/', raw_names2$Name),
            to = paste0('../data/geodata/step1/', raw_names2$New_Name))

rm(fnames, p2raw, raw_names)
```

Subset the attributes of Geodata and Provide standardized naming for the seed zones

```{r}


shapes <- list.files(path = '../data/geodata/step1/', pattern = '.shp$')
fnames <- paste0('../data/geodata/step1/', shapes)

Allium_acuminatum_WUS <- st_read(fnames[1], quiet = T) %>% 
  select(Zone = GRIDCODE)

Artemisia_tridentata_spp_vaseyana_WUS <- st_read(fnames[2], quiet = T) %>% 
  select(Zone = GRIDCODE)

Artemisia_tridentata_spp_tridentata <- st_read(fnames[3], quiet = T) %>% 
  select(Zone = Seedzone)

Artemisia_tridentata_spp_wyomingensis <- st_read(fnames[3], quiet = T) %>% 
  select(Zone = Seedzone)

Astragalus_lonchocarpus_CP <- st_read(fnames[4], quiet = T) %>% 
  select(Zone = OBJECTID)

Bouteloua_gracilis_CP <- st_read(fnames[6], quiet = T) %>% 
  select(Zone = OBJECTID)

Bromus_sitchensis_ssp_carinatus_BM <- st_read(fnames[7], quiet = T) %>% 
  select(Zone = GRIDCODE)
  
Cleome_lutea_CP <- st_read(fnames[8], quiet = T) %>% 
  select(Zone = Axis)

Cleome_serrulata <- st_read(fnames[9], quiet = T) %>% 
  select(Zone = assignment)

Elymus_elymoides_CP <- st_read(fnames[10], quiet = T) %>% 
  select(Zone = OBJECTID)

Elymus_elymoides_WUS <- st_read(fnames[11], quiet = T) %>% 
  select(Zone = GRIDCODE)

Elymus_glaucus_BM <- st_read(fnames[12], quiet = T) %>% 
  select(Zone = ZONE)

Ephedra_nevadensis_MD <- st_read(fnames[13], quiet = T) %>% 
  select(Zone = GRIDCODE)

Eriocoma_hymenoides_WUS <- st_read(fnames[14], quiet = T) %>% 
  select(Zone = GRIDCODE)

Heliomeris_multiflora_CP <- st_read(fnames[15], quiet = T) %>% 
  select(Zone = assignment)

Koeleria_macrantha_CB_GB <- st_read(fnames[16], quiet = T) %>% 
  select(Zone = OBJECTID)

Leymus_cinereus_CB_GB <- st_read(fnames[17], quiet = T) %>% 
  select(Zone = GRIDCODE)

Machaeranthera_canescens_CP <- st_read(fnames[18], quiet = T) %>% 
  select(Zone = Axis)

Pleuraphis_jamesii_CP_GB_MD_WP <- st_read(fnames[19], quiet = T) %>% 
  select(Zone = gridcode)

Pleuraphis_jamesii_CP_GB_MD_WP <- st_read(fnames[19], quiet = T) %>% 
  select(Zone = gridcode)


plot(Leymus_cinereus_CB_GB)
```

