---
title: "Consolidate the formatting of Western Empirical Seed Zones"
author: "steppe"
date: "2025-3-30"
output: html_document
---

Let's use eSTZwritR to try and standardize all of the existing seed transfer zones. 
*big gulp*

```{r, message = F}
# devtools::install_github('sagesteppe/eSTZwritR')
library(eSTZwritR)
library(sf)
library(units)
library(tidyverse)
set.seed(12) # guess I'm going to be the be all end all on the proposed method for ordering
# seed zones. hmmmm.
```

```{r}
dir_out <- '../RetroactivelyReclassified'
```

Mostly we just need to change a few variables... Here is a wrapper function which allows us to do that. 
We'll run it in two modes, one for just moving the data into format, the other for retroactively reclassifying the old data sets based on the aridity index. 
I don't recommend that people start using the old data sets with reclassified zone numbers, but I'll put it out there for the sake of consistency. 

```{r}
wrapper <- function(x, SeedZone, reclass, nrcs_code, sci_name, estz_type, caption, ls = FALSE, city_reduce_no,  ... ){
  
  if(missing(reclass)){reclass <- TRUE}
  if(missing(city_reduce_no)){city_reduce_no<-20}
  if(typeof(x)!='list'){
    taxon <- sf::st_read(x)
  } else {taxon <- x}
  rc <- regionCoding(taxon)
  
  taxon <- sf::st_transform(taxon, 5070)
  taxon <- fieldsmakR(taxon, SeedZone = SeedZone, ...)

  if(reclass){
      ord2 <- orderZones(taxon, n = 5000) 
      taxon <- ord2$Reclassified
  }

  dirmakR(
    outpath = dir_out, sci_name = sci_name, x = taxon, 
    nrcs_code = nrcs_code, regioncode = rc$SuggestedName, overwrite = TRUE,
    estz_type = estz_type) 
  
  taxon <- sf::st_transform(taxon, 4326)
  mapmakR(x = taxon, sci_name = gsub('_', ' ', sci_name), SZName = SZName, landscape = ls,
          city_reduce_no = city_reduce_no,
        outpath = file.path(dir_out, paste0(sci_name, '_', estz_type, 'STZ'), 'Information'), 
         caption = caption)
}
```

```{r indentify all products}
fnames <- paste0('../data/geodata/step1/', 
                 list.files(path = '../data/geodata/step1/', pattern = '.shp$'))
```

## Reclassify products 

Here we retroactively apply the ordered seed zone number approach to the old data sets. 
Note that ordering seed zones this way is not even a core recommendation for new data sets. 
Hopefully folks read a README before they start using this data set. 

```{r acth7}
acth7 <- sf::st_read(file.path(
  system.file(package="eSTZwritR"), "extdata", 'ACTH7.gpkg')
) |>
  select(-area_ha, -GRIDCODE)

wrapper(acth7, SeedZone = 'zone', sci_name = 'Eriocoma_thurberiana', 
        nrcs_code = 'ACTH7',
        estz_type = 'lg', 
        caption = 'Data from Johnson et al. 2017. https://doi.org/10.1016/j.rama.2017.01.004')
```

```{r allium acuminatum}
wrapper(x = st_read(fnames[[grep('Allium_acuminatum', fnames)]]),  
        SeedZone = 'GRIDCODE', SZName = 'ZONE', nrcs_code = 'ALAC4',
        sci_name = 'Allium_acuminatum', 
        estz_type = 'cg', 
        caption = 'Data from Johnson et al. 2013. dx.doi.org/10.1139/cjb-2013-0046')
```

```{r Artemisia}
wrapper(x = fnames[[grep('spp._vaseyana', fnames)]],
        SeedZone = 'GRIDCODE', nrcs_code = 'ARTRV',
        sci_name = 'Artemisia_tridentata_spp_vaseyana', ls = FALSE,
        estz_type = 'cm', city_reduce_no = 30,
        caption = 'Data from Richardson, B. A., & Chaney, L. (2018). https://doi.org/10.1002/eap.1804')

wrapper(
  x = st_read(fnames[[grep('Artemisia_tridentata-WUS', fnames)]]) |> 
    select(SeedZone = Seedzone),
        SeedZone = 'SeedZone', nrcs_code = 'ARTR',
        sci_name = 'Artemisia_tridentata', ls = FALSE,
        estz_type = 'cm', 
        caption = 'Data from Richardson, B. A., & Chaney, L. (2018). https://doi.org/10.1002/eap.1804')
```

```{r Astragalus lonchocarpus}
wrapper(
  x = st_read(fnames[[grep('Astragalus_lonchocarpus-CP', fnames)]]) |> select(-Shape_Leng, -Shape_Area),
        SeedZone = 'OBJECTID', nrcs_code = 'ASLO3',
        sci_name = 'Astragalus_lonchocarpus', 
        estz_type = 'cm', 
        caption = 'Data from Doherty et al. 2017.')

aslo3 <- st_read(fnames[[grep('Astragalus_lonchocarpus-GSTZ', fnames)]]) |>
  rename_with(~ gsub("TS", "BIO4", .x))

wrapper(aslo3, SeedZone = 'assignment', sci_name = 'Astragalus_lonchocarpus', 
        nrcs_code = 'ASLO3', ls = TRUE,
        estz_type = 'lg', 
        caption = 'Data from Massatti & Winkler 2022. https://doi.org/10.1111/2041-210X.13902')

rm(aslo3)
```

```{r bouteloua gracilis}
wrapper(
  x = st_read(fnames[[grep('Bouteloua_gracilis', fnames)]]) |>
    dplyr::select(-c(Shape_Area, Shape_Leng)),
        SeedZone = 'OBJECTID', nrcs_code = 'BOGR2',
        sci_name = 'Bouteloua_gracilis', 
        estz_type = 'cm', 
        caption = 'Data from Doherty et al. 2017.')
```

```{r bromus sitchensis}
wrapper(
  x = st_read(fnames[[grep('Bromus_sitchensis_ssp._carinatus', fnames)]]) |>
    select(-Area, -Perimeter),
        SeedZone = 'GRIDCODE', nrcs_code = 'BOGR2',
        sci_name = 'Bromus_sitchensis_ssp_carinatus', 
        estz_type = 'cg', 
        caption = 'Data from Johnson et al. 2010. doi:10.1139/B10-047')
```

```{r cleome lutea}
cllu2 <- st_read(fnames[[grep('Cleome_lutea', fnames)]])
cllu2 <- rename_with(cllu2, ~ gsub("TS", "BIO4", .x))

wrapper(cllu2, SeedZone = 'Axis', sci_name = 'Cleomella_lutea', 
        nrcs_code = 'CLLU2',
        estz_type = 'lg', 
        caption = 'Data from Massatti et al. 2020. doi: 10.1111/rec.13142')

rm(cllu2)
```

```{r cleome serrulata}
clse <- st_read(fnames[[grep('Cleome_serrulata', fnames)]])
clse <- rename_with(clse, ~ gsub("TS", "BIO4", .x))

wrapper(clse, SeedZone = 'assignment', sci_name = 'Cleome_serrulata', 
        nrcs_code = 'CLSE',
        estz_type = 'lg', caption = 'Data from Massatti 2022. 10.5066/P9H9M79K')
rm(clse)
```

```{r elymus elymoides climate matched}
elel5 <- st_read(fnames[[grep('Elymus_elymoides-CP', fnames)]]) |>
  select(-Shape_Area, -Shape_Leng)

wrapper(elel5, SeedZone = 'OBJECTID', sci_name = 'Elymus_elymoides', 
        nrcs_code = 'ELEL5',
        estz_type = 'cm', caption = 'Data from Doherty et al. 2017.')

elel5 <- st_read(fnames[[grep('Elymus_elymoides-WUS', fnames)]])

wrapper(elel5, SeedZone = 'GRIDCODE', sci_name = 'Elymus_elymoides', 
        nrcs_code = 'ELEL5',
        estz_type = 'cg', caption = 'WHERE DID THIS COME FROM??? WHY DID I START THIS PROJECT? WHY AM I HERE?')
rm(elel5)
```

```{r elymus glaucus}
taxon <- st_read(fnames[[grep('Elymus_glaucus', fnames)]]) |>
  select(-c(ACCURACY, DATA_SOURC, GLOBALID, NAME, REV_DATE, SHAPE_AREA, SHAPE_LEN))

wrapper(taxon, SeedZone = 'OBJECTID', SZName = 'ZONE',
        nrcs_code = 'ELGL', estz_type = 'lg', sci_name = 'Elymus_glaucus', 
        caption = 'Data from Erickson et al. 2004. https://doi.org/10.1139/b04-141')
rm(taxon)
```

```{r Ephedra nevadensis}
wrapper(x = fnames[[grep('Ephedra_nevadensis', fnames)]],
        SeedZone = 'GRIDCODE', nrcs_code = 'EPNE',
        sci_name = 'Ephedra_nevadensis',
        estz_type = 'lg', 
        caption = 'Data from Shyrock et al. 2017. https://doi.org/10.1002/eap.1447')
```

```{r eriocoma hymenoides}
wrapper(x = st_read(fnames[[grep('Eriocoma_hymenoides', fnames)]]),  
        SeedZone = 'GRIDCODE',  SZName = 'ZONE', nrcs_code = 'ACHY',
        sci_name = 'Eriocoma_hymenoides', ls = TRUE,
        estz_type = 'cg', 
        caption = 'Data from Johnson et al. 2012. https://doi.org/10.2111/REM-D-11-00165.1')
```

```{r heliomeris multiflora}
hemu3 <- st_read(fnames[[grep('Heliomeris_multiflora', fnames)]])
hemu3 <- rename_with(hemu3, ~ gsub("PS", "BIO15", .x)) |>
  rename_with(~ gsub("MP", "BIO12", .x))

wrapper(hemu3, SeedZone = 'assignment', sci_name = 'Heliomeris_multiflora', 
        nrcs_code = 'HEMU3',
        estz_type = 'cm', caption = 'Data from Doherty et al. 2017.')
rm(hemu3)
```

```{r koeleria macrantha}
wrapper(x = fnames[[grep('Koeleria_macrantha', fnames)]],
        SeedZone = 'OBJECTID', nrcs_code = 'KOMA',
        sci_name = 'Koeleria_macrantha', 
        estz_type = 'cm', 
        caption = 'Data from Doherty et al. 2017.')
```

```{r leymus cinereus}
wrapper(
  x = st_read(fnames[[grep('Leymus_cinereus', fnames)]]) |> dplyr::select(-c(area_ha, zone)),
  SeedZone = 'GRIDCODE', nrcs_code = 'LECI4',
  sci_name = 'Leymus_cinereus', 
  estz_type = 'cg', 
  caption = 'Data from Johnson & Vance-Borland 2016. https://doi.org/10.1371/journal.pone.0148982')
```

```{r dieteria canescens}
maca <- st_read(fnames[[grep('Machaeranthera_canescens', fnames)]])
maca <- rename_with(maca, ~ gsub("TS", "BIO4", .x)) |>
  rename_with(~ gsub("MDR", "BIO2", .x))
wrapper(maca, SeedZone = 'Axis', sci_name = 'Dieteria_canescens', 
        nrcs_code = 'MACA2',
        estz_type = 'cm', caption = 'Data from Doherty et al. 2017.')
rm(maca)
```

```{r Pleuraphis jamesii}
plja <- st_read(fnames[[grep('Pleuraphis_jamesii', fnames)]])
plja <- rename_with(plja, ~ gsub("Tseas", "BIO4", .x)) |>
  rename_with( ~ gsub('MAT', 'BIO1', .x)) |>
  rename_with( ~ gsub('MDR', 'BIO2', .x)) |>
  select(-population)

wrapper(plja, SeedZone = 'gridcode', sci_name = 'Pleuraphis_jamesii', 
        nrcs_code = 'PLJA', ls = TRUE,
        estz_type = 'cm', caption = 'Data from Massatii 2020')
rm(plja)
```

```{r poa secunda}
pose <- st_read(fnames[[grep('Poa_secunda', fnames)]]) |>
  select(-Area)

wrapper(x = pose, 
        SeedZone = 'GRIDCODE', SZName = 'ZONE', sci_name = 'Poa_secunda', 
        nrcs_code = 'POSE',
        estz_type = 'cg', caption = 'Data from Johnson et al. 2015.')

rm(pose)
```

```{r pseudoroegneria spicata}
pssp6 <- st_read(fnames[[grep('Pseudoroegneria_spicata', fnames)]]) |>
  select(-AREA, -PERIMETER)

wrapper(pssp6, SeedZone = 'GRID_CODE', SZName = 'ZONE', sci_name = 'Pseudoroegneria_spicata', 
        nrcs_code = 'PSSP6', ls = TRUE,
        estz_type = 'cg', 
        caption = 'Data from Bradley St. Clair et al. 2013. doi:10.1111/eva.12077')

rm(pssp6)
```

```{r Sphaeralcea ambigua}
wrapper(x = fnames[[grep('Sphaeralcea_ambigua', fnames)]],
        SeedZone = 'GRIDCODE', nrcs_code = 'SPAM',
        sci_name = 'Sphaeralcea_ambigua', 
        estz_type = 'lg', 
        caption = 'Data from Shyrock et al. 2017. https://doi.org/10.1002/eap.1447')
```

```{r Sphaeralcea parvifolia surfaces}
sppa2 <- st_read(fnames[[grep('Sphaeralcea_parvifolia', fnames)]])
sppa2 <- rename_with(sppa2, ~ gsub("MDR", "BIO2", .x)) |>
  rename_with(~ gsub("Tseas", "BIO4", .x))|>
  rename_with(~ gsub("TwetQ", "BIO8", .x))|>
  rename_with(~ gsub("TwetQ", "BIO8", .x))|>
  rename_with(~ gsub("Pseas", "BIO15", .x))

wrapper(sppa2, SeedZone = 'gridcode', sci_name = 'Sphaeralcea_parvifolia', 
        nrcs_code = 'SPPA2',
        estz_type = 'cm', caption = 'Data from Doherty et al. 2017.')

rm(sppa2)
```

```{r sporobolus cryptandrus}
spcr <- st_read(fnames[[grep('Sporobolus_cryptandrus', fnames)]])
spcr <- rename_with(spcr, ~ gsub("Tseas", "BIO4", .x)) 
wrapper(spcr, SeedZone = 'gridcode', sci_name = 'Sporobolus_cryptandrus', 
        nrcs_code = 'SPCR',
        estz_type = 'cm', caption = 'Data from Doherty et al. 2017.')
rm(spcr)
```

## Combine all vector data to form a multi-layer geopackage

```{r combine vector data into single geopackage}
p <- '../RetroactivelyReclassified'
files <- list.files(p, recursive = TRUE, pattern = '.shp$')
nms <- gsub('[.]shp$', '', basename(files))

# first we create the file 
st_write(
  sf::st_transform(
    sf::st_read(file.path(p, files[1]), quiet = TRUE), 4326), 
  dsn = '../eSTZs_RetroactivelyReclassified.gpkg', 
  layer = nms[1], 
  append = FALSE
)

# and then we simply append each vector data set onto it. 
for (f in seq(from = 2, to = length(files))){
  sf::st_write(
    sf::st_transform(
      sf::st_read(file.path(p, files[f]), quiet = TRUE), 4326),
    dsn = "../eSTZs_RetroactivelyReclassified.gpkg",
    layer = nms[f],
    quiet = TRUE,
    append = TRUE)
}

rm(p, files, nms, f)
```

## Update products 

```{r}
dir_out <- '../RetroactivelyStandardized'
```

```{r acth7}
acth7 <- sf::st_read(file.path(
  system.file(package="eSTZwritR"), "extdata", 'ACTH7.gpkg')
) |>
  select(-area_ha, -GRIDCODE)

wrapper(acth7, SeedZone = 'zone', sci_name = 'Eriocoma_thurberiana', 
        nrcs_code = 'ACTH7', reclass = TRUE,
        estz_type = 'lg', 
        caption = 'Data from Johnson et al. 2017. https://doi.org/10.1016/j.rama.2017.01.004')
```

```{r allium acuminatum}
wrapper(x = st_read(fnames[[grep('Allium_acuminatum', fnames)]]),  
        SeedZone = 'GRIDCODE', SZName = 'ZONE', nrcs_code = 'ALAC4',
        sci_name = 'Allium_acuminatum', reclass = TRUE,
        estz_type = 'cg', 
        caption = 'Data from Johnson et al. 2013. dx.doi.org/10.1139/cjb-2013-0046')
```

```{r Artemisia}
wrapper(x = fnames[[grep('spp._vaseyana', fnames)]],
        SeedZone = 'GRIDCODE', nrcs_code = 'ARTRV', reclass = TRUE,
        sci_name = 'Artemisia_tridentata_spp_vaseyana', ls = FALSE,
        estz_type = 'cm', city_reduce_no = 30,
        caption = 'Data from Richardson, B. A., & Chaney, L. (2018). https://doi.org/10.1002/eap.1804')

wrapper(
  x = st_read(fnames[[grep('Artemisia_tridentata-WUS', fnames)]]) |> 
    select(SeedZone = Seedzone),
        SeedZone = 'SeedZone', nrcs_code = 'ARTR',
        sci_name = 'Artemisia_tridentata', ls = FALSE,
        estz_type = 'cm', reclass = TRUE,
        caption = 'Data from Richardson, B. A., & Chaney, L. (2018). https://doi.org/10.1002/eap.1804')
```

```{r Astragalus lonchocarpus}
wrapper(
  x = st_read(fnames[[grep('Astragalus_lonchocarpus-CP', fnames)]]) |> select(-Shape_Leng, -Shape_Area),
        SeedZone = 'OBJECTID', nrcs_code = 'ASLO3',
        sci_name = 'Astragalus_lonchocarpus', 
        estz_type = 'cm', reclass = TRUE,
        caption = 'Data from Doherty et al. 2017.')

aslo3 <- st_read(fnames[[grep('Astragalus_lonchocarpus-GSTZ', fnames)]]) |>
  rename_with(~ gsub("TS", "BIO4", .x))

wrapper(aslo3, SeedZone = 'assignment', sci_name = 'Astragalus_lonchocarpus', 
        nrcs_code = 'ASLO3', ls = TRUE,
        estz_type = 'lg', reclass = TRUE,
        caption = 'Data from Massatti & Winkler 2022. https://doi.org/10.1111/2041-210X.13902')

rm(aslo3)
```

```{r bouteloua gracilis}
wrapper(
  x = st_read(fnames[[grep('Bouteloua_gracilis', fnames)]]) |>
    dplyr::select(-c(Shape_Area, Shape_Leng)),
        SeedZone = 'OBJECTID', nrcs_code = 'BOGR2',
        sci_name = 'Bouteloua_gracilis', 
        estz_type = 'cm', reclass = TRUE,
        caption = 'Data from Doherty et al. 2017.')
```

```{r bromus sitchensis}
wrapper(
  x = st_read(fnames[[grep('Bromus_sitchensis_ssp._carinatus', fnames)]]) |>
    select(-Area, -Perimeter),
        SeedZone = 'GRIDCODE', nrcs_code = 'BOGR2',
        sci_name = 'Bromus_sitchensis_ssp_carinatus', 
        estz_type = 'cg', reclass = TRUE,
        caption = 'Data from Johnson et al. 2010. doi:10.1139/B10-047')
```


```{r cleome lutea}
cllu2 <- st_read(fnames[[grep('Cleome_lutea', fnames)]])
cllu2 <- rename_with(cllu2, ~ gsub("TS", "BIO4", .x))

wrapper(cllu2, SeedZone = 'Axis', sci_name = 'Cleomella_lutea', 
        nrcs_code = 'CLLU2',
        estz_type = 'lg', reclass = TRUE,
        caption = 'Data from Massatti et al. 2020. doi: 10.1111/rec.13142')

rm(cllu2)
```

```{r cleome serrulata}
clse <- st_read(fnames[[grep('Cleome_serrulata', fnames)]])
clse <- rename_with(clse, ~ gsub("TS", "BIO4", .x))

wrapper(clse, SeedZone = 'assignment', sci_name = 'Cleome_serrulata', 
        nrcs_code = 'CLSE',reclass = TRUE,
        estz_type = 'lg', caption = 'Data from Massatti 2022. 10.5066/P9H9M79K')
rm(clse)
```

```{r elymus elymoides climate matched}
elel5 <- st_read(fnames[[grep('Elymus_elymoides-CP', fnames)]]) |>
  select(-Shape_Area, -Shape_Leng)

wrapper(elel5, SeedZone = 'OBJECTID', sci_name = 'Elymus_elymoides', 
        nrcs_code = 'ELEL5', reclass = TRUE,
        estz_type = 'cm', caption = 'Data from Doherty et al. 2017.')

elel5 <- st_read(fnames[[grep('Elymus_elymoides-WUS', fnames)]])

wrapper(elel5, SeedZone = 'GRIDCODE', sci_name = 'Elymus_elymoides', 
        nrcs_code = 'ELEL5', reclass = TRUE,
        estz_type = 'cg', caption = 'WHERE DID THIS COME FROM??? WHY DID I START THIS PROJECT? WHY AM I HERE?')
rm(elel5)
```

```{r elymus glaucus}
taxon <- st_read(fnames[[grep('Elymus_glaucus', fnames)]]) |>
  select(-c(ACCURACY, DATA_SOURC, GLOBALID, NAME, REV_DATE, SHAPE_AREA, SHAPE_LEN))

wrapper(taxon, SeedZone = 'OBJECTID', SZName = 'ZONE', reclass = TRUE,
        nrcs_code = 'ELGL', estz_type = 'lg', sci_name = 'Elymus_glaucus', 
        caption = 'Data from Erickson et al. 2004. https://doi.org/10.1139/b04-141')
rm(taxon)
```

```{r Ephedra nevadensis}
wrapper(x = fnames[[grep('Ephedra_nevadensis', fnames)]],
        SeedZone = 'GRIDCODE', nrcs_code = 'EPNE',
        sci_name = 'Ephedra_nevadensis',
        estz_type = 'lg', reclass = TRUE,
        caption = 'Data from Shyrock et al. 2017. https://doi.org/10.1002/eap.1447')
```

```{r eriocoma hymenoides}
wrapper(x = st_read(fnames[[grep('Eriocoma_hymenoides', fnames)]]),  
        SeedZone = 'GRIDCODE',  SZName = 'ZONE', nrcs_code = 'ACHY',
        sci_name = 'Eriocoma_hymenoides', ls = TRUE,
        estz_type = 'cg', reclass = TRUE,
        caption = 'Data from Johnson et al. 2012. https://doi.org/10.2111/REM-D-11-00165.1')
```

```{r heliomeris multiflora}
hemu3 <- st_read(fnames[[grep('Heliomeris_multiflora', fnames)]])
hemu3 <- rename_with(hemu3, ~ gsub("PS", "BIO15", .x)) |>
  rename_with(~ gsub("MP", "BIO12", .x))

wrapper(hemu3, SeedZone = 'assignment', sci_name = 'Heliomeris_multiflora', 
        nrcs_code = 'HEMU3', reclass = TRUE,
        estz_type = 'cm', caption = 'Data from Doherty et al. 2017.')
rm(hemu3)
```

```{r koeleria macrantha}
wrapper(x = fnames[[grep('Koeleria_macrantha', fnames)]],
        SeedZone = 'OBJECTID', nrcs_code = 'KOMA',
        sci_name = 'Koeleria_macrantha', 
        estz_type = 'cm', reclass = TRUE,
        caption = 'Data from Doherty et al. 2017.')
```

```{r leymus cinereus}
wrapper(
  x = st_read(fnames[[grep('Leymus_cinereus', fnames)]]) |> dplyr::select(-c(area_ha, zone)),
  SeedZone = 'GRIDCODE', nrcs_code = 'LECI4',
  sci_name = 'Leymus_cinereus', 
  estz_type = 'cg', reclass = TRUE,
  caption = 'Data from Johnson & Vance-Borland 2016. https://doi.org/10.1371/journal.pone.0148982')
```

```{r dieteria canescens}
maca <- st_read(fnames[[grep('Machaeranthera_canescens', fnames)]])
maca <- rename_with(maca, ~ gsub("TS", "BIO4", .x)) |>
  rename_with(~ gsub("MDR", "BIO2", .x))
wrapper(maca, SeedZone = 'Axis', sci_name = 'Dieteria_canescens', 
        nrcs_code = 'MACA2', reclass = TRUE,
        estz_type = 'cm', caption = 'Data from Doherty et al. 2017.')
rm(maca)
```

```{r Pleuraphis jamesii}
plja <- st_read(fnames[[grep('Pleuraphis_jamesii', fnames)]])
plja <- rename_with(plja, ~ gsub("Tseas", "BIO4", .x)) |>
  rename_with( ~ gsub('MAT', 'BIO1', .x)) |>
  rename_with( ~ gsub('MDR', 'BIO2', .x)) |>
  select(-population)

wrapper(plja, SeedZone = 'gridcode', sci_name = 'Pleuraphis_jamesii', 
        nrcs_code = 'PLJA', ls = TRUE, reclass = TRUE,
        estz_type = 'cm', caption = 'Data from Massatii 2020')
rm(plja)
```

```{r poa secunda}
pose <- st_read(fnames[[grep('Poa_secunda', fnames)]]) |>
  select(-Area)

wrapper(x = pose, 
        SeedZone = 'GRIDCODE', SZName = 'ZONE', sci_name = 'Poa_secunda', 
        nrcs_code = 'POSE', reclass = TRUE,
        estz_type = 'cg', caption = 'Data from Johnson et al. 2015.')

rm(pose)
```

```{r pseudoroegneria spicata}
pssp6 <- st_read(fnames[[grep('Pseudoroegneria_spicata', fnames)]]) |>
  select(-AREA, -PERIMETER)

wrapper(pssp6, SeedZone = 'GRID_CODE', SZName = 'ZONE', sci_name = 'Pseudoroegneria_spicata', 
        nrcs_code = 'PSSP6', ls = TRUE,
        estz_type = 'cg', reclass = TRUE,
        caption = 'Data from Bradley St. Clair et al. 2013. doi:10.1111/eva.12077')

rm(pssp6)
```

```{r Sphaeralcea ambigua}
wrapper(x = fnames[[grep('Sphaeralcea_ambigua', fnames)]],
        SeedZone = 'GRIDCODE', nrcs_code = 'SPAM',
        sci_name = 'Sphaeralcea_ambigua', 
        estz_type = 'lg', reclass = TRUE,
        caption = 'Data from Shyrock et al. 2017. https://doi.org/10.1002/eap.1447')
```

```{r Sphaeralcea parvifolia surfaces}
sppa2 <- st_read(fnames[[grep('Sphaeralcea_parvifolia', fnames)]])
sppa2 <- rename_with(sppa2, ~ gsub("MDR", "BIO2", .x)) |>
  rename_with(~ gsub("Tseas", "BIO4", .x))|>
  rename_with(~ gsub("TwetQ", "BIO8", .x))|>
  rename_with(~ gsub("TwetQ", "BIO8", .x))|>
  rename_with(~ gsub("Pseas", "BIO15", .x))

wrapper(sppa2, SeedZone = 'gridcode', sci_name = 'Sphaeralcea_parvifolia', 
        nrcs_code = 'SPPA2', reclass = TRUE,
        estz_type = 'cm', caption = 'Data from Doherty et al. 2017.')

rm(sppa2)
```

```{r sporobolus cryptandrus}
spcr <- st_read(fnames[[grep('Sporobolus_cryptandrus', fnames)]])
spcr <- rename_with(spcr, ~ gsub("Tseas", "BIO4", .x)) 
wrapper(spcr, SeedZone = 'gridcode', sci_name = 'Sporobolus_cryptandrus', 
        nrcs_code = 'SPCR', reclass = TRUE,
        estz_type = 'cm', caption = 'Data from Doherty et al. 2017.')
rm(spcr)
```

## Combine all vector data to form a multi-layer geopackage

```{r combine vector data into single geopackage}
p <- '../RetroactivelyStandardized'
files <- list.files(p, recursive = TRUE, pattern = '.shp$')
nms <- gsub('[.]shp$', '', basename(files))

# first we create the file 
st_write(
  sf::st_transform(sf::st_read(file.path(p, files[1])), 4326), 
  dsn = '../RetroactivelyStandardized.gpkg', 
  layer = nms[1], 
  append = FALSE
)

# and then we simply append each vector data set onto it. 
for (f in seq(from = 2, to = length(files))){
  sf::st_write(
    sf::st_transform(
      sf::st_read(file.path(p, files[f]), quiet = TRUE), 4326),
    dsn = "../RetroactivelyStandardized.gpkg",
    layer = nms[f],
    quiet = TRUE,
    append = TRUE)
}

rm(p, files, nms, f)
```

```{r}

```

