---
title: "Consolidate the formatting of Western Empirical Seed Zones"
author: "steppe"
date: "2025-3-30"
output: html_document
---

Let's use eSTZwritR to try and standardize all of the existing seed transfer zones. 
*big gulp*

```{r, message = F}
# devtools::install_github('sagesteppe/eSTZwritR')
library(eSTZwritR)
library(sf)
library(units)
library(tidyverse)
set.seed(12) # guess I'm going to be the be all end all on the proposed method for ordering
# seed zones. hmmmm.
```

```{r}
dir_out <- '../RetroactivelyStandardized'
```

Mostly we just need to change a few variables... Here is a wrapper function which allows us to do that. 
```{r}
wrapper <- function(x, SeedZone, nrcs_code, sci_name, estz_type, caption, ls = FALSE, ... ){
  
  if(typeof(x)!='list'){
    taxon <- sf::st_read(x)
  } else {taxon <- x}
  rc <- regionCoding(taxon)
  
  taxon <- sf::st_transform(taxon, 5070)
  taxon <- fieldsmakR(taxon, SeedZone = SeedZone, ...)

  ord2 <- orderZones(taxon, n = 3000) 
  taxon <- ord2$Reclassified
  return(taxon)

  dirmakR(
    outpath = dir_out, sci_name = sci_name, x = taxon, 
    nrcs_code = nrcs_code, regioncode = rc$SuggestedName, overwrite = TRUE,
    estz_type = estz_type) 
  
  taxon <- sf::st_transform(taxon, 4326)
  mapmakR(x = taxon, sci_name = gsub('_', ' ', sci_name), SZName = SZName, landscape = ls,
        outpath = file.path(dir_out, paste0(sci_name, '_', estz_type, 'STZ'), 'Information'), 
         caption = caption)
  return(taxon)
}
```

Start with the example for the package
```{r acth7, eval = F}
acth7 <- sf::st_read(file.path(
  system.file(package="eSTZwritR"), "extdata", 'ACTH7.gpkg')
)
rc <- regionCoding(acth7)

acth7$area_ha <- set_units(set_units(acth7$area_ha, 'ha'), 'acres')
acth7 <- fieldsmakR(acth7, SeedZone = 'zone', ID = 'ID', AreaAcres = 'area_ha')
acth7 <- dplyr::select(acth7, -c(GRIDCODE))

ord2 <- orderZones(acth7, SeedZone, n = 1000) 
acht7 <- ord2$Reclassified

dirmakR(
  outpath = dir_out, sci_name = 'Eriocoma thurberiana', x = acth7, 
  nrcs_code = 'ACTH7', regioncode = rc$SuggestedName, 
  estz_type = 'lg') 

mapmakR(x = acth7, sci_name = 'Eriocoma thurberiana', SZName = SZName, landscape = FALSE, 
        outpath = file.path(dir_out, 'Eriocoma_thurberiana_lgSTZ', 'Information'), 
         caption = 'Data from Johnson et al. 2017. https://doi.org/10.1016/j.rama.2017.01.004')
rm(acth7)
```

identify all products which need renaming. 

```{r indentify all products}
fnames <- paste0('../data/geodata/step1/', 
                 list.files(path = '../data/geodata/step1/', pattern = '.shp$'))
```


```{r allium acuminatum}
wrapper(x = st_read(fnames[[grep('Allium_acuminatum', fnames)]]),  
        SeedZone = 'GRIDCODE', SZName = 'ZONE', nrcs_code = 'ALAC4',
        sci_name = 'Allium_acuminatum', 
        estz_type = 'cg', 
        caption = 'Data from Johnson et al. 2013. dx.doi.org/10.1139/cjb-2013-0046')
```

NOTE ACHY Is a weird one... the gridcode is one thing, and the SZNames are numbers which do not match (they are consecutive version of gridcode). 
We will overwrite the SZName to make them confluent with our 'new' values. To do this we just omit the original zone column. 

```{r achy, eval = F}

wrapper(x = st_read(fnames[[grep('Eriocoma_hymenoides', fnames)]]) |> select(-area_ha, -zone),  
        SeedZone = 'GRIDCODE',  nrcs_code = 'LECI4',
        sci_name = 'Eriocoma_hymenoides', 
        estz_type = 'cg', 
        caption = 'Data from Johnson et al. 2012. https://doi.org/10.2111/REM-D-11-00165.1')

```

The formatting of the sage files are unique, we will come back to them. 
```{r Artemisia, eval = F}

wrapper(x = fnames[[grep('spp._vaseyana', fnames)]],
        SeedZone = 'GRIDCODE', nrcs_code = 'ARTRV',
        sci_name = 'Artemisia_tridentata_spp_vaseyana', ls = FALSE,
        estz_type = 'cm', 
        caption = 'Data from Richardson, B. A., & Chaney, L. (2018). https://doi.org/10.1002/eap.1804')

```


```{r leymus cinereus, eval = F}
wrapper(
  x = st_read(fnames[[grep('Leymus_cinereus', fnames)]]) |> dplyr::select(-c(area_ha, zone)),
  SeedZone = 'GRIDCODE', nrcs_code = 'LECI4',
  sci_name = 'Leymus_cinereus', 
  estz_type = 'cg', 
  caption = 'Data from Johnson & Vance-Borland 2016. https://doi.org/10.1371/journal.pone.0148982')

```

```{r ASLO BOGR2 KOMA SPCR HEMU3 SPPA surfaces, eval = F}

wrapper(
  x = st_read(fnames[[grep('Astragalus_lonchocarpus-CP', fnames)]]) |> select(-Shape_Leng, -Shape_Area),
        SeedZone = 'OBJECTID', nrcs_code = 'ASLO3',
        sci_name = 'Astragalus_lonchocarpus', 
        estz_type = 'cm', 
        caption = 'Data from Doherty et al. 2017.')

wrapper(x = fnames[[grep('Bouteloua_gracilis', fnames)]],
        SeedZone = 'OBJECTID', nrcs_code = 'BOGR2',
        sci_name = 'Bouteloua_gracilis', 
        estz_type = 'cm', 
        caption = 'Data from Doherty et al. 2017.')

wrapper(x = fnames[[grep('Koeleria_macrantha', fnames)]],
        SeedZone = 'OBJECTID', nrcs_code = 'KOMA',
        sci_name = 'Koeleria_macrantha', 
        estz_type = 'cm', 
        caption = 'Data from Doherty et al. 2017.')

maca <- st_read(fnames[[grep('Machaeranthera_canescens', fnames)]])
maca <- rename_with(maca, ~ gsub("TS", "BIO4", .x)) |>
  rename_with(~ gsub("MDR", "BIO2", .x))
wrapper(maca, SeedZone = 'Axis', sci_name = 'Dieteria_canescens', 
        nrcs_code = 'MACA2',
        estz_type = 'cm', caption = 'Data from Doherty et al. 2017.')

spcr <- st_read(fnames[[grep('Sporobolus_cryptandrus', fnames)]])
spcr <- rename_with(spcr, ~ gsub("Tseas", "BIO4", .x)) 
wrapper(spcr, SeedZone = 'gridcode', sci_name = 'Sporobolus_cryptandrus', 
        nrcs_code = 'SPCR',
        estz_type = 'cm', caption = 'Data from Doherty et al. 2017.')

hemu3 <- st_read(fnames[[grep('Heliomeris_multiflora', fnames)]])
hemu3 <- rename_with(hemu3, ~ gsub("PS", "BIO15", .x)) |>
  rename_with(~ gsub("MP", "BIO12", .x))

wrapper(hemu3, SeedZone = 'assignment', sci_name = 'Heliomeris_multiflora', 
        nrcs_code = 'HEMU3',
        estz_type = 'cm', caption = 'Data from Doherty et al. 2017.')

sppa2 <- st_read(fnames[[grep('Sphaeralcea_parvifolia', fnames)]])
sppa2 <- rename_with(sppa2, ~ gsub("MDR", "BIO2", .x)) |>
  rename_with(~ gsub("Tseas", "BIO4", .x))|>
  rename_with(~ gsub("TwetQ", "BIO8", .x))|>
  rename_with(~ gsub("TwetQ", "BIO8", .x))|>
  rename_with(~ gsub("Pseas", "BIO15", .x))

wrapper(sppa2, SeedZone = 'gridcode', sci_name = 'Sphaeralcea_parvifolia', 
        nrcs_code = 'SPPA2',
        estz_type = 'cm', caption = 'Data from Doherty et al. 2017.')

rm(hemu3, spcr)
```

```{r poa secunda, eval = F}
pose <- st_read(fnames[[grep('Poa_secunda', fnames)]]) |>
  select(-Area)

# these warnings are 'safe' https://github.com/r-spatial/sf/issues/306
wrapper(x = pose, 
        SeedZone = 'GRIDCODE', SZName = 'ZONE', sci_name = 'Poa_secunda', 
        nrcs_code = 'POSE',
        estz_type = 'cg', caption = 'Data from Johnson et al. 2015.')
```


```{r Pleuraphis jamesii, eval = F}

plja <- st_read(fnames[[grep('Pleuraphis_jamesii', fnames)]])
plja <- rename_with(plja, ~ gsub("Tseas", "BIO4", .x)) |>
  rename_with( ~ gsub('MAT', 'BIO1', .x)) |>
  rename_with( ~ gsub('MDR', 'BIO2', .x)) |>
  select(-population)
wrapper(plja, SeedZone = 'gridcode', sci_name = 'Pleuraphis_jamesii', 
        nrcs_code = 'PLJA', ls = TRUE,
        estz_type = 'cm', caption = 'Data from Massatii 2020')
```

```{r pseudoroegneria spicata}
pssp6 <- st_read(fnames[[grep('Pseudoroegneria_spicata', fnames)]]) |>
  select(-AREA, -PERIMETER)

wrapper(pssp6, SeedZone = 'GRID_CODE', SZName = 'ZONE', sci_name = 'Pseudoroegneria_spicata', 
        nrcs_code = 'PSSP6', ls = TRUE,
        estz_type = 'cg', 
        caption = 'Data from Bradley St. Clair et al. 2013. doi:10.1111/eva.12077')

```


```{r astragalus lonchocarpus genetic, eval = F}

aslo3 <- st_read(fnames[[grep('Astragalus_lonchocarpus-GSTZ', fnames)]]) |>
  rename_with(~ gsub("TS", "BIO4", .x))

wrapper(aslo3, SeedZone = 'assignment', sci_name = 'Astragalus_lonchocarpus', 
        nrcs_code = 'ASLO3', ls = TRUE,
        estz_type = 'lg', 
        caption = 'Data from Massatti & Winkler 2022. https://doi.org/10.1111/2041-210X.13902')
```

```{r cleome lutea, eval = F}

cllu2 <- st_read(fnames[[grep('Cleome_lutea', fnames)]])
colnames(cllu2) <- gsub('TS', 'BIO4', colnames(cllu2))

rc <- regionCoding(cllu2)

cllu2 <- sf::st_transform(cllu2, 5070)
cllu2 <- fieldsmakR(cllu2, SeedZone = 'Axis')

cllu2 <- sf::st_make_valid(cllu2)
ord2 <- orderZones(cllu2, SeedZone, n = 3000) 
cllu2 <- ord2$Reclassified

dirmakR(
    outpath = dir_out, sci_name = 'Cleome_lutea', x = cllu2, 
    nrcs_code = 'CLLU2', regioncode = rc$SuggestedName, overwrite = TRUE,
    estz_type = 'lg') 
  
cllu2 <- sf::st_transform(cllu2, 4326)
mapmakR(x = cllu2, sci_name = gsub('_', ' ', 'Cleome_lutea'), SZName = SZName, landscape = FALSE,
        outpath = file.path(dir_out, paste0('Cleome_lutea', '_', 'lg', 'STZ'), 'Information'), 
         caption = 'Data from Massatti et al. 2020. doi: 10.1111/rec.13142')

rm(cllu2)
```

```{r cleome serrulata eval = F}
clse <- st_read(fnames[[grep('Cleome_serrulata', fnames)]])
colnames(clse) <- gsub('TS', 'BIO4', colnames(clse))

rc <- regionCoding(clse)

clse <- sf::st_transform(clse, 5070)
clse <- fieldsmakR(clse, SeedZone = 'assignment')

ord2 <- orderZones(clse, SeedZone, n = 3000) 
clse <- ord2$Reclassified

dirmakR(
    outpath = dir_out, sci_name = 'Cleome_serrulata', x = clse, 
    nrcs_code = 'CLSE', regioncode = rc$SuggestedName, overwrite = TRUE,
    estz_type = 'lg') 
  
clse <- sf::st_transform(clse, 4326)
mapmakR(x = clse, sci_name = gsub('_', ' ', 'Cleome_serrulata'), SZName = SZName, landscape = FALSE,
        outpath = file.path(dir_out, paste0('Cleome_serrulata', '_', 'lg', 'STZ'), 'Information'), 
         caption = 'Data from Massatti et al. 2020. doi: 10.1111/rec.13142')
```


```{r elymus glaucus}
taxon <- st_read(fnames[[grep('Elymus_glaucus', fnames)]])
rc <- regionCoding(taxon)

taxon <- sf::st_transform(taxon, 5070)
clse <- fieldsmakR(taxon, SeedZone = 'OBJECTID', SZName = 'ZONE')

clse <- dplyr::select(clse, -c(ACCURACY, DATA_SOURC, GLOBALID, NAME, REV_DATE, SHAPE_AREA, SHAPE_LEN))

ord2 <- orderZones(clse, SeedZone, n = 3000) 
clse <- ord2$Reclassified

dirmakR(
    outpath = dir_out, sci_name = 'Elymus_glaucus', x = clse, 
    nrcs_code = 'ELGL', regioncode = rc$SuggestedName, overwrite = TRUE,
    estz_type = 'cg') 
  
clse <- sf::st_transform(clse, 4326)
mapmakR(x = clse, sci_name = gsub('_', ' ', 'Elymus_glaucus'), SZName = SZName, landscape = FALSE,
        outpath = file.path(dir_out, paste0('Elymus_glaucus', '_', 'cg', 'STZ'), 'Information'), 
         caption = 'Data from Erickson et al. 2004. https://doi.org/10.1139/b04-141')
```

```{r Ephedra nevadensis and Sphaeralcea ambigua}
wrapper(x = fnames[[grep('Ephedra_nevadensis', fnames)]],
        SeedZone = 'GRIDCODE', nrcs_code = 'EPNE',
        sci_name = 'Ephedra_nevadensis', 
        estz_type = 'lg', 
        caption = 'Data from Shyrock et al. 2017. https://doi.org/10.1002/eap.1447')

wrapper(x = fnames[[grep('Sphaeralcea_ambigua', fnames)]],
        SeedZone = 'GRIDCODE', nrcs_code = 'SPAM',
        sci_name = 'Sphaeralcea_ambigua', 
        estz_type = 'lg', 
        caption = 'Data from Shyrock et al. 2017. https://doi.org/10.1002/eap.1447')
```





something weird here - revisit it, shouldnt be clipped so close to this area. 

```{r elymus elymoides climate matched}

elel5 <- st_read(fnames[[grep('Elymus_elymoides-CP', fnames)]])

rc <- regionCoding(elel5)

elel5 <- sf::st_transform(elel5, 5070)
elel5 <- fieldsmakR(elel5, SeedZone = 'OBJECTID')

elel5 <- orderZones(elel5, SeedZone, n = 3000) 
elel5 <- ord2$Reclassified

dirmakR(
    outpath = dir_out, sci_name = 'Elymus_elymoides', x = elel5, 
    nrcs_code = 'ELEL5', regioncode = rc$SuggestedName, overwrite = TRUE,
    estz_type = 'cm') 
  
elel5 <- sf::st_transform(elel5, 4326)
mapmakR(x = elel5, sci_name = gsub('_', ' ', 'Elymus_elymoides'), SZName = SZName, landscape = FALSE,
        outpath = file.path(dir_out, paste0('Elymus_elymoides', '_', 'cm', 'STZ'), 'Information'), 
         caption = 'Data from Doherty et al. 2017.')
```

Something weird here too. revisit it

```{r elymus elymoides landscape }

elel5 <- st_read(fnames[[grep('Elymus_elymoides-WUS', fnames)]])

rc <- regionCoding(elel5)

elel5 <- sf::st_transform(elel5, 5070)
elel5 <- fieldsmakR(elel5, SeedZone = 'GRIDCODE')

elel5 <- orderZones(elel5, SeedZone, n = 3000) 
elel5 <- ord2$Reclassified
ggplot() + 
  geom_sf(data = elel5)

dirmakR(
    outpath = dir_out, sci_name = 'Elymus_elymoides', x = elel5, 
    nrcs_code = 'ELEL5', regioncode = rc$SuggestedName, overwrite = TRUE,
    estz_type = 'cg') 
  
elel5 <- sf::st_transform(elel5, 4326)
mapmakR(x = elel5, sci_name = gsub('_', ' ', 'Elymus_elymoides'), SZName = SZName, landscape = FALSE,
        outpath = file.path(dir_out, paste0('Elymus_elymoides', '_', 'cg', 'STZ'), 'Information'), 
         caption = 'Data from Doherty et al. 2017.')
```