---
title: "Making Figures for Poster"
author: "Bri Wieferich and steppe"
date: "2024-07-24"
output: html_document
---

# Import and tidy data. 

We will only use functions contained within R packages in the tidyverse ecosystem. 
The `readxl` package is part of the tidyverse ecosystem, but one of the kind of more
far flung packages which needs to be directly loaded by name rather than in the
`library(tidyverse)` call. 
We'll also use the `sf` package to create some shapes for illustrating our suggestions!
There are certainly some other packages for this, but... 
I think `sf` is one of the all-around best R packages out there, and it can be used
to accomplish just about anything graphics related which is meant to be clear and concise. 
```{r}
library(tidyverse)
library(readxl)
library(sf)
```

```{r load relevant data for plotting}
p2data <- '../data/raw/eSTZ_consistency_formatting.xlsx'
excel_sheets(p2data)
filenames <- read_excel(p2data, sheet = 'Filename schema')
maps <- read_excel(p2data, sheet = 'pdf maps formatted')
```

```{r Refine naming conventions for plotting}

filenames <- filenames |> 
  mutate(Trait = str_remove(Trait, '.*-')) |>

# We don't want to overwhelm the viewer, so we will create barplots of just 
# 3 main traits here, 'Binomial', 'Geographic_Extent', and 'Seed_Zone', 
# we will write the following text in the caption for the plots.

# "File names had three components which were only rarely present, the resolution (3 times), the Coordinate Reference System (1 time), in one instance a file contained information on two species."
  filter(Trait %in% c('Binomial', 'Geographic_Extent', 'Seed_Zone')) 
```


```{r}

# We are going to create a barplot, which we want to use to showcase the current
# disorganization of the existing file names. 

binom_plot <- filenames %>% 
  
  # when we save a plot to an object, by default Rstudio will not render it into 
  # the viewing panel (lower right, 'Plots' panel)
  filter(Trait == 'Binomial') |>
  
    # We are going to convert the character string responses to a 'factor', this
  # will allow us to easily display the values in an order, we will always put the
  # desired file name component at the bottom of the chart. 
  mutate(Response = 
           factor(Response, levels = c('scientific_name', 'NRCS_PLANTS_upper', 
                                       'NRCS_PLANTS_lower', 'common_name'))) %>% 
  
  # ggplot is low-order plotting device. it will require the following arguments: 
  # 1 data = the dataset which will be plotted, in this instance we denote that the 
  # dataset is being 'piped in' from above, and we can explicitly denote that with
  # the value '.' which means LOOK HERE in programming, and can also be seen when
  # using relative paths to load data. 
  # ggplot then uses the 'aes' or aesthetics arguments to determine HOW it will plot
  # data from the data argument. These Variable (column) names should NEVER be
  # quoted, and must be spelled exactly. 
  # within aes 'y' determines which variable will be plotted along the Y axis
  # within aes 'x' determines which variables will be plotted along the X axis 
  ggplot(., aes(y = Value, x = '', fill = forcats::fct_rev(Response))) + 
  
  # ggplot works on the notions of 'GEOMS', basically what a geom is a different 
  # type of plot. Geoms exist for nearly all common forms of data visualization 
  # if you want to learn more about the arguments  to geom_bar we can use the 
  # '?' function  to get some data on it... '?geom_bar()'
  geom_bar(position = "stack", stat = "identity") + 
  
  # labs this is a function we can use to specify labels for the plot parameters
  labs(
    x = NULL,
    y = 'Count', 
    title = 'How do filenames\nrepresent the species?', 
    fill = 'Character State'
      ) + 
  
  # ggplot has a few main default themes for styling the plot in GENERAL, e.g. 
  # whether there are hash lines on the background or not.. 
  theme_minimal() + 
  
  # we are going to label the portion of the chart, and place the legend beneath
  # rather than to the side of the plot. 
  geom_text(aes(label = Response), position = position_stack(vjust = 0.5)) +
  theme(legend.position="bottom", 
        plot.title = element_text(hjust = 0.5),
        legend.title.position = 'top', 
        legend.title = element_text(hjust = 0.5)) +

  # by default ggplot puts a lot of white space between the content of the plot
  # and the axis labels; we are going to reduce the x-axis spacing. 
  scale_x_discrete(expand = c(0,0)) +
  
  # we are also going to select some preferable colours to the default scheme
  scale_fill_manual(values = c('#C0C5C1', '#F08080', '#6279B8',  '#496F5D'))


# We will repeat the process with Geographic extent. 
extent_plot <- filenames %>% 
  filter(Trait == 'Geographic_Extent') |>
  mutate(Response = 
           factor(Response, levels = c('Specified', 'Not-specified'))) %>% 
  ggplot(., aes(y = Value, x = '', fill = forcats::fct_rev(Response))) + 
  geom_bar(position = "stack", stat = "identity") + 
  labs(y = 'Count', x = NULL, title = 'Is the geographic\nextent specified?', 
       fill = 'Specified') +
  theme_minimal() +
  theme(
    legend.position="bottom", 
    plot.title = element_text(hjust = 0.5),
    legend.title.position = 'top', 
    legend.title = element_text(hjust = 0.5)
    ) + 
  scale_x_discrete(expand = c(0,0)) +
  
  geom_text(aes(label = Response), position = position_stack(vjust = 0.5)) +
  scale_fill_manual(values = c('#C0C5C1',  '#496F5D')) 


file_plot <- filenames %>% 
  filter(Trait == 'Seed_Zone') |>
  mutate(Response = 
           factor(Response, levels = c('zone', 'seed_zone', 'sz', 'missing'))) %>% 
  ggplot(., aes(y = Value, x = '', fill = forcats::fct_rev(Response))) + 
  geom_bar(position = "stack", stat = "identity", width=1) + 
  labs(y = 'Count', x = NULL, title = 'Do files mention\nthey contain STZs?', 
       fill = 'Seed Zone Specified') + 
    scale_fill_manual(values = c('#C0C5C1', '#F08080', '#6279B8',  '#496F5D'))+
  theme_minimal() +
  theme(
    legend.position="bottom", 
    plot.title = element_text(hjust = 0.5),
    legend.title.position = 'top', 
    legend.title = element_text(hjust = 0.5)
    ) + 
  scale_x_discrete(expand = c(0,0)) +
  geom_text(aes(label = Response), position = position_stack(vjust = 0.5)) 


fnames <- binom_plot  + file_plot + extent_plot + 
  plot_annotation(title = 'Inconsistency in File Naming', 
                  theme = theme(plot.title = element_text(size = 18, hjust = 0.5))) 

ggsave('../poster/filenames.png', fnames, width = 10, height = 10, units = 'in')

rm(binom_plot, file_plot, extent_plot, filenames, fnames)
```

Maps, many map features take on binary outcomes. 
We can display them as 'Yes/No'. 
We'll order the bars from the most to least "No's"

```{r}

maps <- maps |> 
  mutate(Trait = str_remove(Trait, '.*-')) |>
  filter(Trait %in% c('Title', 'North_Arrow', 'Sources', 'Zone_Key',
                      'Distance_Key', 'State_Lines')) |>
  mutate(Trait = factor(Trait, levels = c('Title', 'North_Arrow', 'Sources', 'Zone_Key',
                      'Distance_Key', 'State_Lines')))

ggplot(data = maps, aes(fill = Response, y = Value, x = Trait)) + 
  geom_bar(position = 'fill', stat = 'identity') + 
  scale_fill_manual(values = c('#00203F','#ADEFD1')) + 
  theme_classic() + 
  labs(title = 'Is the content of maps consistent?', fill = 'Component', 
  y = 'Proportion of Maps with element') +
  theme(plot.title = element_text(hjust = 0.5)) + 
  scale_x_discrete(labels=c('Title', 'North Arrow', 'Sources', 'Zone Legend',
                      'Distance Key', 'State Lines')) 
  
ggsave('../poster/maps.png', width = 10, height = 10, units = 'in')
```



# Create an example filename

```{r}

fname
```


# Make folder directories

We will use sf to update our directory tree. 
We want this to be a nice big catchy tree that helps to pull attendees in!
We'll use continuous colours, moving from darker to lighter, to try and pull this off. 

```{r Create Directory Tree Example}

# we will use the simple features package to create polygon shapes which we can then
# map both text, and lines onto. 

# We will create many rhombic polygons to showcase the various levels of the directory (folders)
# we'll then 'buffer', or add a predefined distance around  these polygons. The
# real reason we want to buffer the polygons is so that we can create 'soft corners'
# These are generally seen as more gentle and inviting than hard corners. 
# We achieve this effect via the combination of the joinStyle and endCapStyle 
# arguments, with a very minimal buffer distance (1)

# sf objects can take on just about any projection in existence, however by not specifying
# any real datum, we are just gonna keep everything in a cartesian grid. 

# notice that despite us creating square-like shapes we have to define 
# 5 pairs of coordinates (x, y), this is because the 5th set of coordinates
# needs to go back to the origin point (the 1st set of coordinates) to close 
# the polygon.
Genus_epithet_lgSTZ <- sf::st_buffer(
  st_polygon(list(rbind(c(0,0), c(10,0), c(10,10), c(0,10), c(0,0)))),
  dist = 1,
  joinStyle = 'ROUND', 
  endCapStyle = "ROUND") 

Data <- sf::st_buffer(
  st_polygon(list(rbind(c(1,5.5), c(1,9), c(9,9), c(9,5.5), c(1,5.5)))),
  dist = 1,
  joinStyle = 'ROUND', 
  endCapStyle = "ROUND") 

Information <- sf::st_buffer(
  st_polygon(list(rbind(c(1,1), c(1,3), c(9,3), c(9,1), c(1,1)))),
  dist = 1,
  joinStyle = 'ROUND', endCapStyle = "ROUND")

Raster <- sf::st_buffer(
  st_polygon(list(rbind(c(2,8.5), c(2,8.5), c(8.75,8.5), c(8.5,8.5), c(2,8.5)))),
  dist = 1,
  joinStyle = 'ROUND', endCapStyle = "ROUND")

Vector <- sf::st_buffer(
  st_polygon(list(rbind(c(2,6), c(2,6), c(8.75,6), c(7,6), c(2,6)))),
  dist = 1,
  joinStyle = 'ROUND', endCapStyle = "ROUND") 

# notice that here we are creating a 'point' type geometry, and placing it
# into a dataframe, We can map the 'Label' to a function later to type this out. 
# all of these labels will be combined with the function geom_sf_label so that we
# have a nice label box around them. 
ge_lg_lab <- data.frame(Label = 'Genus_epithet_lgSTZ', x = 1, y = 11) |>
  st_as_sf(coords = c(x = 'x', y = 'y'))
data_lab <- data.frame(Label = 'Data', x = 0.5, y = 10) |>
  st_as_sf(coords = c(x = 'x', y = 'y'))
raster_lab <- data.frame(Label = 'Raster', x = 2, y = 9.25) |>
  st_as_sf(coords = c(x = 'x', y = 'y'))
vector_lab <- data.frame(Label = 'Vector', x = 2, y = 7) |>
  st_as_sf(coords = c(x = 'x', y = 'y'))
information_lab <- data.frame(Label = 'Information', x = 1.25, y = 4) |>
  st_as_sf(coords = c(x = 'x', y = 'y'))

# These are text data again, and we will simply write them onto the plot using
# geom_sf_text. 
raster_file <- data.frame(Label = 'ABCD_lgSTZ_UCB-LCB.tif', x = 6.25, y = 8.5) |>
  st_as_sf(coords = c(x = 'x', y = 'y')) 
vector_file <- data.frame(Label = 'ABCD_lgSTZ_UCB-LCB.shp', x = 6.25, y = 6) |>
  st_as_sf(coords = c(x = 'x', y = 'y')) 
citation <- data.frame(Label = 'citation.txt', x = 2.5, y = 3) |>
  st_as_sf(coords = c(x = 'x', y = 'y')) 
map <- data.frame(Label = 'map.txt', x = 2.2, y = 2) |>
  st_as_sf(coords = c(x = 'x', y = 'y')) 
paper <- data.frame(Label = 'Paper_about_making_the_STZ.pdf', x = 4.9, y = 1) |>
  st_as_sf(coords = c(x = 'x', y = 'y')) 

ggplot () +
  geom_sf(data = Genus_epithet_lgSTZ, fill = NA) + 
  geom_sf(data = Data) + 
  geom_sf(data = Information) + 
  geom_sf(data = Raster) + 
  geom_sf(data = Vector) +

  # we can draw some 'line' type geometries straight onto the plot, this means
  # we don't have to delete (rm) as many objects at the end of this r chunk. 
  geom_sf(data = st_as_sfc("LINESTRING(-0.5 11,-0.5 10, 0 10)")) + # upper bend to 'Data'
  geom_sf(data = st_as_sfc("LINESTRING(0.5 10, 0.5 9.25, 1.5 9.25)")) + # bend to 'Raster'
  geom_sf(data = st_as_sfc("LINESTRING(0.5 10, 0.5 7, 1.5 7)")) + # bend to 'Vector'
  geom_sf(data = st_as_sfc("LINESTRING(2 9, 2 8.5, 3 8.5)")) + # Raster file contents
  geom_sf(data = st_as_sfc("LINESTRING(2 7, 2 6, 3 6)")) + # Vector file contents
  geom_sf(data = st_as_sfc("LINESTRING(-0.5 10, -0.5 4, 0 4)")) +  # bend to 'Information' 
  geom_sf(data = st_as_sfc("LINESTRING(0.5 4, 0.5 3, 1.25 3)")) +  # bends throughout 'Information 1
  geom_sf(data = st_as_sfc("LINESTRING(0.5 3, 0.5 2, 1.25 2)")) +  # bends throughout 'Information 2
  geom_sf(data = st_as_sfc("LINESTRING(0.5 2, 0.5 1, 1.25 1 )")) + # bends throughout 'Information 2

  # add on our sub directory names
  geom_sf_label(data = ge_lg_lab, aes(label = Label)) +
  geom_sf_label(data = data_lab, aes(label = Label)) +
  geom_sf_label(data = raster_lab, aes(label = Label)) +
  geom_sf_label(data = vector_lab, aes(label = Label)) +
  geom_sf_label(data = information_lab, aes(label = Label)) +
  
  # throw in the example file names. 
  geom_sf_text(data = raster_file, aes(label = Label)) + 
  geom_sf_text(data = vector_file, aes(label = Label)) +
  geom_sf_text(data = citation, aes(label = Label)) +
  geom_sf_text(data = map, aes(label = Label)) + 
  geom_sf_text(data = paper, aes(label = Label)) +
  theme_void()

ggsave('../poster/ExampleDirectory.png', width = 5, height = 5, units = 'in', dpi = 300)


rm(Genus_epithet_lgSTZ, Data, Information, Raster, Vector, 
   ge_lg_lab, data_lab, raster_lab, vector_lab, information_lab,
   raster_file, vector_file, citation, map, paper)
```

