---
title: "Making Figures for Poster"
author: "Bri Wieferich and steppe"
date: "2024-07-24"
output: html_document
---

# Import and tidy data. 

We will only use functions contained within R packages in the tidyverse ecosystem. 
The `readxl` package is part of the tidyverse ecosystem, but one of the kind of more
far flung packages which needs to be directly loaded by name rather than in the
`library(tidyverse)` call. 
We'll also use the `sf` package to create some shapes for illustrating our suggestions!
There are certainly some other packages for this, but... 
I think `sf` is one of the all-around best R packages out there, and it can be used
to accomplish just about anything graphics related which is meant to be clear and concise. 

```{r}
library(tidyverse)
library(ggtext)
library(readxl)
library(sf)
library(showtext)
library(patchwork)
library(gtable)
```

```{r load relevant data for plotting}
p2data <- '../data/raw/eSTZ_consistency_formatting.xlsx'
excel_sheets(p2data)
filenames <- read_excel(p2data, sheet = 'Filename schema')
maps <- read_excel(p2data, sheet = 'pdf maps formatted')
```

```{r Refine naming conventions for plotting}

filenames <- filenames |> 
  mutate(Trait = str_remove(Trait, '.*-')) |>

# We don't want to overwhelm the viewer, so we will create barplots of just 
# 3 main traits here, 'Binomial', 'Geographic_Extent', and 'Seed_Zone', 
# we will write the following text in the caption for the plots.

# "File names had three components which were only rarely present, the resolution (3 times), the Coordinate Reference System (1 time), in one instance a file contained information on two species."
  filter(Trait %in% c('Binomial', 'Geographic_Extent', 'Seed_Zone')) 
```


```{r}

# We are going to create a barplot, which we want to use to showcase the current
# disorganization of the existing file names. 

showtext_auto() # helps ggsave avoid shrinking text = 
showtext_opts(dpi = 300)

binom_plot <- filenames %>% 
  
  # when we save a plot to an object, by default Rstudio will not render it into 
  # the viewing panel (lower right, 'Plots' panel)
  filter(Trait == 'Binomial') |>
  
    # We are going to convert the character string responses to a 'factor', this
  # will allow us to easily display the values in an order, we will always put the
  # desired file name component at the bottom of the chart. 
  mutate(Response = 
           factor(Response, levels = c('NRCS_PLANTS_upper', 'NRCS_PLANTS_lower',
                                       'scientific_name',  'common_name'))) %>% 
  
  # ggplot is low-order plotting device. it will require the following arguments: 
  # 1 data = the dataset which will be plotted, in this instance we denote that the 
  # dataset is being 'piped in' from above, and we can explicitly denote that with
  # the value '.' which means LOOK HERE in programming, and can also be seen when
  # using relative paths to load data. 
  # ggplot then uses the 'aes' or aesthetics arguments to determine HOW it will plot
  # data from the data argument. These Variable (column) names should NEVER be
  # quoted, and must be spelled exactly. 
  # within aes 'y' determines which variable will be plotted along the Y axis
  # within aes 'x' determines which variables will be plotted along the X axis 
  ggplot(., aes(y = Value, x = '', fill = forcats::fct_rev(Response))) + 
  
  # ggplot works on the notions of 'GEOMS', basically what a geom is a different 
  # type of plot. Geoms exist for nearly all common forms of data visualization 
  # if you want to learn more about the arguments  to geom_bar we can use the 
  # '?' function  to get some data on it... '?geom_bar()'
  geom_bar(position = "stack", stat = "identity") + 
  
  # labs this is a function we can use to specify labels for the plot parameters
  labs(
    x = NULL,
    y = 'Count', 
    title = 'How do filenames\nrepresent the species?', 
    fill = 'Character State'
      ) + 
  
  # ggplot has a few main default themes for styling the plot in GENERAL, e.g. 
  # whether there are hash lines on the background or not.. 
  theme_minimal() + 
  
  # we are going to label the portion of the chart, and place the legend beneath
  # rather than to the side of the plot. 
  geom_text(aes(label = Response), position = position_stack(vjust = 0.5)) +
  theme(
    panel.grid.major = element_blank(),
    panel.border = element_blank(),
    rect = element_rect(fill = "transparent"),
    panel.background = element_rect(fill='transparent'),
    legend.position="none", 
    plot.title = element_text(hjust = 0.5),
    legend.title.position = 'top', 
    legend.title = element_text(hjust = 0.5)
) +

  # by default ggplot puts a lot of white space between the content of the plot
  # and the axis labels; we are going to reduce the x-axis spacing. 
  scale_x_discrete(expand = c(0,0)) +
  
  # we are also going to select some preferable colours to the default scheme
  scale_fill_manual(values = c('#EF476F', '#E9D985', '#007991', '#439A86'))


# We will repeat the process with Geographic extent. 
extent_plot <- filenames %>% 
  filter(Trait == 'Geographic_Extent') |>
  mutate(Response = if_else(Response == 'Specified', 'Yes', 'No')) |>
  mutate(Response = 
           factor(Response, levels = c('Yes', 'No'))) %>% 
  ggplot(., aes(y = Value, x = '', fill = forcats::fct_rev(Response))) + 
  geom_bar(position = "stack", stat = "identity") + 
  labs(y = 'Count', x = NULL, title = 'Is the geographic\nextent specified?', 
       fill = 'Specified') +
  theme_minimal() +
  theme(
    panel.grid.major = element_blank(),
    panel.border = element_blank(),
    rect = element_rect(fill = "transparent"),
    panel.background = element_rect(fill='transparent'),
    legend.position="none", 
    plot.title = element_text(hjust = 0.5),
    legend.title.position = 'top', 
    legend.title = element_text(hjust = 0.5), 
    axis.text.y = element_blank()
    ) + 
  labs(y = NULL) + 
  scale_x_discrete(expand = c(0,0)) +
  
  geom_text(aes(label = Response), position = position_stack(vjust = 0.5)) +
  scale_fill_manual(values = c('#EF476F',  '#439A86')) 


file_plot <- filenames %>% 
  filter(Trait == 'Seed_Zone') |>
  mutate(Response = 
           factor(Response, levels = c('zone', 'seed_zone', 'sz', 'missing'))) %>% 
  ggplot(., aes(y = Value, x = '', fill = forcats::fct_rev(Response))) + 
  geom_bar(position = "stack", stat = "identity", width=1) + 
  labs(y = 'Count', x = NULL, title = 'Do files mention\nthey contain STZs\n if so how?', 
       fill = 'Seed Zone Specified') + 
    scale_fill_manual(values = c('#EF476F', '#BCD8C1', '#E9D985',  '#439A86'))+
  theme_minimal() +
  theme(
    panel.grid.major = element_blank(),
    panel.border = element_blank(),
    rect = element_rect(fill = "transparent"),
    panel.background = element_rect(fill='transparent'),
    legend.position="none", 
    plot.title = element_text(hjust = 0.5),
    legend.title.position = 'top', 
    legend.title = element_text(hjust = 0.5),
    axis.text.y = element_blank()
    ) + 
  labs(y = NULL) + 
  scale_x_discrete(expand = c(0,0)) +
  geom_text(aes(label = Response), position = position_stack(vjust = 0.5)) 



fnames <- binom_plot  + file_plot + extent_plot + 
  plot_annotation(
    title = 'Do file names have consistent components?', 
    theme = theme(
      plot.title = element_text(size = 18, hjust = 0.5),
    rect = element_rect(fill = "transparent"),
    panel.background = element_rect(fill='transparent')))

fnames
ggsave('../poster/filenames.png', fnames, bg = 'transparent')

rm(binom_plot, file_plot, extent_plot, filenames, fnames)
```

Maps, many map features take on binary outcomes. 
We can display them as 'Yes/No'. 
We'll order the bars from the most to least "No's"

```{r}

maps <- maps |> 
  mutate(Trait = str_remove(Trait, '.*-')) |>
  filter(Trait %in% c('Title', 'North_Arrow', 'Sources', 'Zone_Key',
                      'Distance_Key', 'State_Lines')) |>
  mutate(Trait = factor(Trait, levels = c('Title', 'North_Arrow', 'Sources', 'Zone_Key',
                      'Distance_Key', 'State_Lines')))

ggplot(data = maps, aes(fill = Response, y = Value, x = Trait)) + 
  geom_bar(position = 'fill', stat = 'identity') + 
  scale_fill_manual(values = c('#EF476F', '#007991')) + 
  theme_classic() + 
  labs(title = 'Do maps have consistent content?', fill = 'Component', 
  y = 'Proportion of Maps with element') +
  theme(
    rect = element_rect(fill = "transparent"),
    panel.background = element_rect(fill='transparent'),
    axis.text = element_text(size = 16),
    axis.title  = element_text(size = 18, color='black'),
    legend.text = element_text(size = 18),
    legend.title = element_text(size = 20),
    legend.position = c(0.901, 0.2), 
    legend.margin=margin(c(1,10,10,10)),
    legend.key = element_rect(fill = "grey70"),
    plot.title = element_text(size = 24, hjust = 0.5)
    ) + 
  scale_x_discrete(labels=c('Title', 'North Arrow', 'Sources', 'Zone Legend',
                      'Distance Key', 'State Lines')) 
  
ggsave('../poster/maps.png', width = 10, height = 10, units = 'in', bg = 'transparent')
```


# Make folder directories

We will use sf to update our directory tree. 
We want this to be a nice big catchy tree that helps to pull attendees in!
We'll use continuous colours, moving from darker to lighter, to try and pull this off. 

```{r Create Directory Tree Example}

# we will use the simple features package to create polygon shapes which we can then
# map both text, and lines onto. 

# We will create many rhombic polygons to showcase the various levels of the directory (folders)
# we'll then 'buffer', or add a predefined distance around  these polygons. The
# real reason we want to buffer the polygons is so that we can create 'soft corners'
# These are generally seen as more gentle and inviting than hard corners. 
# We achieve this effect via the combination of the joinStyle and endCapStyle 
# arguments, with a very minimal buffer distance (1)

# sf objects can take on just about any projection in existence, however by not specifying
# any real datum, we are just gonna keep everything in a cartesian grid. 

# notice that despite us creating square-like shapes we have to define 
# 5 pairs of coordinates (x, y), this is because the 5th set of coordinates
# needs to go back to the origin point (the 1st set of coordinates) to close 
# the polygon.
Genus_epithet_lgSTZ <- sf::st_buffer(
  st_polygon(list(rbind(c(0,0), c(10,0), c(10,10), c(0,10), c(0,0)))),
  dist = 1,
  joinStyle = 'ROUND', 
  endCapStyle = "ROUND") 

Data <- sf::st_buffer(
  st_polygon(list(rbind(c(1,5.5), c(1,9), c(9,9), c(9,5.5), c(1,5.5)))),
  dist = 1,
  joinStyle = 'ROUND', 
  endCapStyle = "ROUND") 

Information <- sf::st_buffer(
  st_polygon(list(rbind(c(1,1), c(1,3), c(9,3), c(9,1), c(1,1)))),
  dist = 1,
  joinStyle = 'ROUND', endCapStyle = "ROUND")

Raster <- sf::st_buffer(
  st_polygon(list(rbind(c(2,8.5), c(2,8.5), c(8.75,8.5), c(8.5,8.5), c(2,8.5)))),
  dist = 1,
  joinStyle = 'ROUND', endCapStyle = "ROUND")

Vector <- sf::st_buffer(
  st_polygon(list(rbind(c(2,6), c(2,6), c(8.75,6), c(7,6), c(2,6)))),
  dist = 1,
  joinStyle = 'ROUND', endCapStyle = "ROUND") 

# notice that here we are creating a 'point' type geometry, and placing it
# into a dataframe, We can map the 'Label' to a function later to type this out. 
# all of these labels will be combined with the function geom_sf_label so that we
# have a nice label box around them. 
ge_lg_lab <- data.frame(Label = 'Genus_epithet_lgSTZ', x = 1, y = 11) |>
  st_as_sf(coords = c(x = 'x', y = 'y'))
data_lab <- data.frame(Label = 'Data', x = 0.5, y = 10) |>
  st_as_sf(coords = c(x = 'x', y = 'y'))
raster_lab <- data.frame(Label = 'Raster', x = 2, y = 9.25) |>
  st_as_sf(coords = c(x = 'x', y = 'y'))
vector_lab <- data.frame(Label = 'Vector', x = 2, y = 7) |>
  st_as_sf(coords = c(x = 'x', y = 'y'))
information_lab <- data.frame(Label = 'Information', x = 1.25, y = 4) |>
  st_as_sf(coords = c(x = 'x', y = 'y'))

# These are text data again, and we will simply write them onto the plot using
# geom_sf_text. 
raster_file <- data.frame(Label = 'ABCD_lgSTZ_UCB-LCB.tif', x = 6.25, y = 8.5) |>
  st_as_sf(coords = c(x = 'x', y = 'y')) 
vector_file <- data.frame(Label = 'ABCD_lgSTZ_UCB-LCB.shp', x = 6.25, y = 6) |>
  st_as_sf(coords = c(x = 'x', y = 'y')) 
citation <- data.frame(Label = 'citation.txt', x = 2.5, y = 3) |>
  st_as_sf(coords = c(x = 'x', y = 'y')) 
map <- data.frame(Label = 'map.txt', x = 2.2, y = 2) |>
  st_as_sf(coords = c(x = 'x', y = 'y')) 
paper <- data.frame(Label = 'Paper_about_making_the_STZ.pdf', x = 4.75, y = 1) |>
  st_as_sf(coords = c(x = 'x', y = 'y')) 

ggplot () +
  geom_sf(data = Genus_epithet_lgSTZ, fill = '#BCD8C1') + 
  geom_sf(data = Data, fill = '#439A86') + 
  geom_sf(data = Information, fill = '#439A86') + 
  geom_sf(data = Raster, fill = '#007991') + 
  geom_sf(data = Vector, fill = '#007991') +

  # we can draw some 'line' type geometries straight onto the plot, this means
  # we don't have to delete (rm) as many objects at the end of this r chunk. 
  geom_sf(data = st_as_sfc("LINESTRING(-0.5 11,-0.5 10, 0 10)")) + # upper bend to 'Data'
  geom_sf(data = st_as_sfc("LINESTRING(0.5 10, 0.5 9.25, 1.5 9.25)")) + # bend to 'Raster'
  geom_sf(data = st_as_sfc("LINESTRING(0.5 10, 0.5 7, 1.5 7)")) + # bend to 'Vector'
  geom_sf(data = st_as_sfc("LINESTRING(2 9, 2 8.5, 3 8.5)")) + # Raster file contents
  geom_sf(data = st_as_sfc("LINESTRING(2 7, 2 6, 3 6)")) + # Vector file contents
  geom_sf(data = st_as_sfc("LINESTRING(-0.5 10, -0.5 4, 1 4)")) +  # bend to 'Information' 
  geom_sf(data = st_as_sfc("LINESTRING(0.5 4, 0.5 3, 1.25 3)")) +  # bends throughout 'Information 1
  geom_sf(data = st_as_sfc("LINESTRING(0.5 3, 0.5 2, 1.25 2)")) +  # bends throughout 'Information 2
  geom_sf(data = st_as_sfc("LINESTRING(0.5 2, 0.5 1, 1.25 1 )")) + # bends throughout 'Information 2

  # add on our sub directory names
  geom_sf_label(data = ge_lg_lab, aes(label = Label), fill = '#E9D985') +
  geom_sf_label(data = data_lab, aes(label = Label), fill = '#E9D985') +
  geom_sf_label(data = raster_lab, aes(label = Label), fill = '#E9D985') +
  geom_sf_label(data = vector_lab, aes(label = Label), fill = '#E9D985') +
  geom_sf_label(data = information_lab, aes(label = Label), fill = '#E9D985') +
  
  # throw in the example file names. 
  geom_sf_text(data = raster_file, aes(label = Label)) + 
  geom_sf_text(data = vector_file, aes(label = Label)) +
  geom_sf_text(data = citation, aes(label = Label)) +
  geom_sf_text(data = map, aes(label = Label)) + 
  geom_sf_text(data = paper, aes(label = Label)) +
  theme_void()

ggsave('../poster/ExampleDirectory.png', width = 5, height = 5, units = 'in', dpi = 300)

rm(Genus_epithet_lgSTZ, Data, Information, Raster, Vector, 
   ge_lg_lab, data_lab, raster_lab, vector_lab, information_lab,
   raster_file, vector_file, citation, map, paper)
```


# Create an example filename

```{r}

fname <- data.frame(Label = 'ACTH7_gSTZ_UCB-LCB.ext', x = 0, y = 0) |>
  st_as_sf(coords = c(x = 'x', y = 'y'))
region <- 'Up to the two DOI regions which the STZ overlaps the most\n 
these can be identified using the `regionCoding` function from the R package.' 

nrcs_poly <- st_polygon(
  list(
    rbind(
      c(-0.0105,-0.0013), c(-0.0105,0.01), c(-0.035, 0.01), 
      c(-0.035, 0.0025), c(-0.025, 0.0024), 
      c(-0.025,-0.0013), c(-0.0105,-0.0013)
      )
    )
  ) |>
  st_buffer(dist = 0.0005,
  joinStyle = 'ROUND', 
  endCapStyle = "ROUND") 

stz_poly <- st_polygon(
  list(
    rbind(
      c(-0.0095, 0.0018), c(0.00015, 0.0018), c(0.00015, -0.015), c(-0.025, -0.015), 
      c(-0.025, -0.0023), c(-0.0095, -0.0023), c(-0.0095, 0.0018) 
      )
    )
  )|>
  st_buffer(dist = 0.0005,
  joinStyle = 'ROUND', 
  endCapStyle = "ROUND") 

region_poly <- st_polygon(
  list( 
    rbind( 
      c(-0.0095, 0.0027), c(0.001, 0.0027), c(0.001, -0.002), c(0.0165, -0.002),
      c(0.0165, 0.015), c(-0.0095, 0.015),
      c(-0.0095, 0.0027) 
      )
    )
  )|>
  st_buffer(dist = 0.0005,
  joinStyle = 'ROUND', 
  endCapStyle = "ROUND") 

ext_poly <- st_polygon(
  list( 
    rbind( 
      c(0.0175, -0.002), c(0.0175, 0.0018), c(0.025, 0.0018), c(0.025, -0.01), 
      c(0.001, -0.01), c(0.001, -0.003), c(0.0175, -0.003), 
      c(0.0175, -0.002)
      )
    )
  )|>
  st_buffer(dist = 0.0005,
  joinStyle = 'ROUND', 
  endCapStyle = "ROUND") 


NRCS <- data.frame(Label = 'NRCS PLANTS code from\nhttps://plants.usda.gov/', x = -0.0225, y = 0.0065) |>
  st_as_sf(coords = c(x = 'x', y = 'y')) 

STZ <- data.frame(
  Label =  'The type of data/analysis used\nto develop the STZ. One of:\n(g) genetic, (cm) climate matched,\n or (cg) common garden',
                  x = -0.012, y = -0.009) |>
  st_as_sf(coords = c(x = 'x', y = 'y')) 

STZ <- data.frame(
  Label =  'The type of data/analysis used\nto develop the STZ. One of:\n(g) genetic, (cm) climate matched,\n or (cg) common garden',
                  x = -0.012, y = -0.009) |>
  st_as_sf(coords = c(x = 'x', y = 'y')) 

REGION <- data.frame(
  Label = 'Up to two DOI regions which\nthe STZ overlaps the most.', 
  x = 0.004, y = 0.009) |>
  st_as_sf(coords = c(x = 'x', y = 'y')) 

EXT <- data.frame(
  Label = 'File extensions, for rasters `.tif`\nis preferred, for vector data\nshapefiles.', 
  x = 0.0135, y = -0.0065) |>
  st_as_sf(coords = c(x = 'x', y = 'y')) 
  
ggplot() + 
  geom_sf(data = nrcs_poly,  linetype = 2, fill = '#E9D985') +  
  geom_sf(data = stz_poly, linetype = 2, fill = '#007991') +  
  geom_sf(data = region_poly,  linetype = 2, fill = '#BCD8C1') + 
  geom_sf(data = ext_poly, linetype = 2, fill = '#EF476F') + 
  geom_sf_text(data = fname, aes(label = Label), fontface = 'bold', size = 8) +  
  labs(x = NULL, y = NULL, title = 'File naming convention') +  
  theme_void() +
  theme(
    plot.title = element_text(size = 24, hjust = 0.5)) + 
  coord_sf(ylim = c(-0.02, 0.02), xlim = c(-0.035, 0.035)) +
  
  geom_sf_text(data = NRCS, aes(label = Label)) +
  geom_sf_text(data = STZ, aes(label = Label)) +
  geom_sf_text(data = REGION, aes(label = Label)) + 
  geom_sf_text(data = EXT, aes(label = Label))
  

ggsave('../poster/ExampleFileName.png', height = 7.5, width = 7.5, units = 'in', dpi = 300)

# rm(fname, nrcs_poly, stz_poly, region_poly, ext_poly, NRCS, STZ, REGION, EXT)
```



# Analysis of field names

This was an ABSOLUTE MESS and I ended up tabulating by hand because R had no way of dealing with this nonsense. 


```{r}
p <- '../data/geodata/step1'
f <- file.path(p, list.files(p, pattern = '[.]shp$'))

colReporter <- function(x){

  dat <- sf::st_read(x, quiet = T)

  col_head <- head(dat, n = 10) |>
    sf::st_drop_geometry() |>
    cbind(
      Geometry = as.character(unique(sf::st_geometry_type(dat))),
      File = basename(x))

}


out <- lapply(f, colReporter)

# we want to score three things

# 1) Does the Seed Zone containing column contain the word 'Zone'
# 2) Is the Geometry type Polygon or Multipolygon?
# 3) Is area included ? (Y/N) Are the units labelled?

rm(p, f, out, colReporter)
```



```{r}

vals <- data.frame(
  Trait = c(rep(c('Zone', 'Multipolygon', 'Area'), each = 2), 'Area'), 
  Value = c(8, 18, 13, 13, 3, 7, 16), 
  Response = c('Yes', 'No', 'Yes', 'No', 'Yes', 'Yes, but unlabelled', 'No') 
)

showtext_auto() # helps ggsave avoid shrinking text = 
showtext_opts(dpi = 300)

  
zones_plot <- vals %>% 
  filter(Trait == 'Zone') %>% 
  ggplot(., aes(y = Value, x = '', fill = Response)) + 
  geom_bar(position = "stack", stat = "identity", width = 1) + 
  labs(y = 'Count', x = NULL, title = 'Is the Seed Zone\nfield labelled?', 
       fill = 'Zone') + 
    scale_fill_manual(values = c('#EF476F',  '#BCD8C1'))+
  theme_minimal() +
  
  geom_text(aes(label = Response), position = position_stack(vjust = 0.5)) + 
  theme(
    legend.position="none", 
    panel.grid.major = element_blank(),
    panel.border = element_blank(),
    plot.title = element_text(hjust = 0.5),
    legend.title.position = 'top', 
    legend.title = element_text(hjust = 0.5), 
    axis.text.y = element_blank(), 
    rect = element_rect(fill = "transparent"),
    panel.background = element_rect(fill='transparent')
    ) + 
  labs(y = NULL) + 
  scale_x_discrete(expand = c(0,0)) 

polygons_plot <- vals %>% 
  filter(Trait == 'Multipolygon') %>% 
  ggplot(., aes(y = Value, x = '', fill = Response)) + 
  geom_bar(position = "stack", stat = "identity", width = 1) + 
  labs(y = 'Count', x = NULL, title = 'Are the geometries\npolygons?', 
       fill = 'Multipolygon') + 
    scale_fill_manual(values = c('#EF476F',  '#BCD8C1'))+
  theme_minimal() +
  theme(
    legend.position="none", 
    panel.grid.major = element_blank(),
    panel.border = element_blank(),
    plot.title = element_text(hjust = 0.5),
    legend.title.position = 'top', 
    legend.title = element_text(hjust = 0.5), 
    rect = element_rect(fill = "transparent"),
    panel.background = element_rect(fill='transparent')
    ) + 
  scale_x_discrete(expand = c(0,0)) +
  scale_y_continuous(breaks = seq(0, 30, by = 5)) + 
  geom_text(aes(label = Response), position = position_stack(vjust = 0.5)) 


areas_plot <- vals %>% 
  filter(Trait == 'Area') %>% 
  mutate(Response = 
           factor(Response, 
                  levels = c('No', 'Yes, but unlabelled', 'Yes'))) %>% 


  ggplot(., aes(y = Value, x = '', fill = Response)) + 
  geom_bar(position = "stack", stat = "identity", width = 1) + 
  labs(y = 'Count', x = NULL, title = 'Is an "Area" field Present-\nand labelled?', 
       fill = 'Multipolygon') + 
    scale_fill_manual(values = c('#EF476F', '#007991', '#BCD8C1'))+
  theme_minimal() +
    scale_x_discrete(expand = c(0,0)) +
  theme(
    legend.position="none", 
    panel.border = element_blank(),
    panel.grid.major = element_blank(),
    plot.title = element_text(hjust = 0.5),
    legend.title.position = 'top', 
    legend.title = element_text(hjust = 0.5), 
    axis.text.y = element_blank(), 
    axis.text = element_text(size = 50),
    rect = element_rect(fill = "transparent"),
    panel.background = element_rect(fill='transparent')
    ) + 
  labs(y = NULL) + 
  geom_text(aes(label = Response), position = position_stack(vjust = 0.5)) 

fields <- polygons_plot + zones_plot + areas_plot + 
  plot_annotation(title = 'Do fields have consistent names across files?', 
                  theme = theme(
                    plot.title = element_text(size = 18, hjust = 0.5), 
                    rect = element_rect(fill = "transparent"),
                    panel.background = element_rect(fill='transparent'))
                  ) 
  
ggsave('../poster/fields.png', fields, bg = 'transparent')
rm(zones_plot, polygons_plot, areas_plot)
```



# Rules for Fields within the file. 


```{r Rules textbox}

df <- tibble(
  label = c(
    'Use UpperCamelCase for field names whenever possible.',
    'The four field names above should be in every file.',
    'Do you need to include extra variables? If so place them "after" these four fields.',
    'If using Bioclim variables, use their\nofficial bioclim designations - e.g. BIO8.',
    'If using a statistic use: min, max, mean, R, SD; prefixed with "_" e.g. "BIO8_R".',
    'Have SeedZone numbers in ascending order from the hottest and driest areas.',
    'Descriptive names (SZName) e.g. "SW montane" are useful additions to the Seed Zones.',
    '"ID" reflects each polygon geometry - if using multipolygons the ID will be "SeedZone".'
  ),
  x = 0,
  y = seq(from = 40, to = 5, by = -5),
  fill = '#BCD8C1',
  orientation = c("upright"),
  color = c("black")
)


ggplot(df) +
  aes(
    x, y, label = label, color = color, fill = fill, box.colour = '#439A86', 
    orientation = orientation
  ) +
  geom_textbox(width = unit(1, "npc")) +
  scale_discrete_identity(aesthetics = c("color", "fill", "orientation")) +
  xlim(-1, 1) + 
  theme_void() + 
  theme(
    panel.background = element_rect(fill = "#007991")
    )

ggsave('../poster/FieldRules.png',  height = 4, units = 'in', dpi = 300)

rm(df)
```


Example df for field names

```{r}

df <- data.frame(
  ID = 1:4,
  SeedZone = 1:4,
  SZName = c('Salt Desert', 'Desert Scrub', 'Pinyon-Juniper/Oak Brush', 'Montane'), 
  AreaAcres = c(12340, 14230, 30142, 9872),
  'BIO1_R' = c(20.2, 19.1, 15.1, 12.3),
  'BIO2_mean' = c(5.1, 7.1, 10.1, 12.3)
)

library(gt)

gt_tbl <- 
  gt(df) |>
  tab_header(
    title = "Example field names in a shapefile",
  ) |>
  tab_options(
    table.background.color = '#BCD8C1', 
    column_labels.background.color = '#E9D985') |>
  tab_style(
    style = cell_fill(color = "#007991" |> adjust_luminance(steps = 2)),
    locations = cells_body(columns = 1:4)
  ) |>
  tab_style(
    style = cell_fill(color = "#439A86" |> adjust_luminance(steps = 2)),
    locations = cells_body(columns = 5:6)
  ) |>
  tab_footnote(
    footnote = "The first four (blue) fields should be in every file. More fields are optional."
  )

gt_tbl |> 
  gtsave('../poster/ExampleFields-table.png', expand = 10)

gtExtras::gtsave_extra(gt_tbl, '../poster/ExampleFields-table.png', expand = 10)
```



